### 2017-04-11 article.tex
Added a TeX stub

\documentclass{article}

\begin{document}

  This is a test!

\end{document}

### 2017-04-12 make_polyglot.sh
Started creating a shell script to make the polyglot

@@ -1,0 +1,60 @@
+if [ "$#" -ne 2 ]; then
+    (>&2 echo "Usage: $0 SOURCE_PDF OUTPUT_PDF\n")
+    exit 1
+fi
+
+SOURCE=$1
+OUTPUT=$2
+
+BRANCH_NAME=PolyglotBranch
+
+git ls-files $OUTPUT --error-unmatch 1>/dev/null 2>&1
+if [ $? -ne 1 ]; then
+    (>&2 echo "Error: The output file $OUTPUT is already tracked by git!\n       This will not work! Please choose another output file.\n")
+    exit 2
+fi
+
+git rev-parse --verify $BRANCH_NAME 1>/dev/null 2>&1
+if [ $? -eq 0 ]; then
+    (>&2 echo "Error: A branch named $BRANCH_NAME already exists!\n       You need to either delete this branch\n       or set a different value for the BRANCH_NAME variable\n       in $0\n")
+    exit 3
+fi
+
+CURRENT_BRANCH=`git rev-parse --abbrev-ref HEAD`
+
+cp $SOURCE $OUTPUT
+
+rm -f ${OUTPUT}.log
+
+git stash save MAKINGPOLYGLOT 1>>${OUTPUT}.log 2>&1
+git stash list | grep -q MAKINGPOLYGLOT
+HAS_STASH=$?
+
+if [ $HAS_STASH -eq 0 ]; then
+    echo "There are local changes! Saving them to a stash so they are not lost..."
+fi
+
+echo "Creating temporary branch ${BRANCH_NAME}..."
+git checkout -b $BRANCH_NAME
+if [ $HAS_STASH -eq 0 ]; then
+    echo "Applying the stash to the new branch..."
+    git stash apply 1>>${OUTPUT}.log 2>&1
+fi
+git update-index --add --cacheinfo 100644 `git hash-object -w $OUTPUT` $OUTPUT
+TREE_HASH=`git write-tree`
+echo 'Polyglot PDF' | git commit-tree $TREE_HASH
+git commit -a -m 'Creating the Polyglot'
+DIR="$( cd "$( dirname "${BASH_SOURCE[0]}" )" && pwd )"
+PATH=$DIR:$PATH git bundle create ${OUTPUT}.bundle --do-not-compress `git hash-object $OUTPUT` --all
+echo "Switching back to branch ${CURRENT_BRANCH}..."
[…]

### 2017-04-12 make_polyglot.sh
The Makefile and shell script are now working

@@ -35,17 +35,18 @@ if [ $HAS_STASH -eq 0 ]; then
 fi
 
 echo "Creating temporary branch ${BRANCH_NAME}..."
-git checkout -b $BRANCH_NAME
+git checkout -b $BRANCH_NAME 1>>${OUTPUT}.log 2>&1
 if [ $HAS_STASH -eq 0 ]; then
     echo "Applying the stash to the new branch..."
     git stash apply 1>>${OUTPUT}.log 2>&1
 fi
-git update-index --add --cacheinfo 100644 `git hash-object -w $OUTPUT` $OUTPUT
+echo "Making the polyglot..."
+git update-index --add --cacheinfo 100644 `git hash-object -w $OUTPUT` $OUTPUT 1>>${OUTPUT}.log 2>&1
 TREE_HASH=`git write-tree`
-echo 'Polyglot PDF' | git commit-tree $TREE_HASH
-git commit -a -m 'Creating the Polyglot'
+echo 'Polyglot PDF' | git commit-tree $TREE_HASH 1>>${OUTPUT}.log 2>&1
+git commit -a -m 'Creating the Polyglot' 1>>${OUTPUT}.log 2>&1
 DIR="$( cd "$( dirname "${BASH_SOURCE[0]}" )" && pwd )"
-PATH=$DIR:$PATH git bundle create ${OUTPUT}.bundle --do-not-compress `git hash-object $OUTPUT` --all
+PATH=$DIR/git:$PATH git bundle create ${OUTPUT}.bundle --do-not-compress `git hash-object $OUTPUT` --all 1>>${OUTPUT}.log 2>&1
 echo "Switching back to branch ${CURRENT_BRANCH}..."
 git checkout $CURRENT_BRANCH 1>>${OUTPUT}.log 2>&1
 echo "Deleting temporary branch ${BRANCH_NAME}..."

### 2017-04-12 article.tex
Started making the PDF content
Background on `git bundle`

@@
 
-  This is a test!
+\title{This PDF is a Git Repo\\Containing its Own \LaTeX\ Source}
+\author{Evan Sultanik}
+\date{April 11, 2017}
+
+\maketitle
+
+Have you ever heard of the \texttt{git bundle} command? I hadn't. It
+bundles a set of Git objects---potentially even an entire
+repository---into a single file. Git allows you to treat that file as
+if it were a standard Git database, so you can do things like clone a
+repo directly from it. Its purpose is to easily sneakernet pushes or
+even whole repositories across air gaps.
 

### 2017-04-12 make_polyglot.sh
Moved the file size test to the script.

@@ -6,6 +6,10 @@ fi
 SOURCE=$1
 OUTPUT=$2
 
+if [ `wc -c <$SOURCE` -gt 65535 ]; then
+    (>&2 echo "Warning: $SOURCE is greater than 0xFFFF bytes. This might not work!")
+fi
+
 BRANCH_NAME=PolyglotBranch
 
 git ls-files $OUTPUT --error-unmatch 1>/dev/null 2>&1

### 2017-04-14 article.tex
Apparently have to make the PDF minor version less than 5 to get Adobe to load it
Started on a file format figure, but that pushed the PDF over 0xFFFF bytes!

@@
 \documentclass{article}
+
+\pdfminorversion=4
+
+\usepackage{tikz}
+\usetikzlibrary{fit}
 
@@
 
-\title{This PDF is a Git Repo\\Containing its Own \LaTeX\ Source}
+\title{This PDF is a Git Repository\\Containing its Own \LaTeX\ Source}
 \author{Evan Sultanik}
@@
 
+The file format for Git bundles doesn't appear to be formally
+specified anywhere, however, inspecting \texttt{bundle.c} reveals that
+it's relatively straightforward:
+\begin{center}
+\begin{tikzpicture}
+  \node[text width=0.9\hsize] at (0,0) (sig) {\footnotesize\texttt{\# v2 git bundle}};
+  \node[inner sep=0mm,fit=(sig),draw,rounded corners,dashed,minimum width=0.9\hsize] (sigbox) {};
+  \node[draw,rounded corners,fill=white] at (sigbox.north) (siglabel) {\scriptsize Git Bundle Signature};
+  \node[text width=0.9\hsize,anchor=north west,yshift=-1mm] at (sig.south west) (b1) {\footnotesize\tt 3aa340a2e3d125ab6703e5c9bdfede2054a9c0c5 refs/heads/master};
+  \node[text width=0.9\hsize,anchor=north west,yshift=-1mm] at (b1.south west) (b2) {\footnotesize\tt 3aa340a2e3d125ab6703e5c9bdfede2054a9c0c5 refs/remotes/origin/master};
+  \node[text width=0.9\hsize,anchor=north west,yshift=-1mm] at (b2.south west) (b3) {\footnotesize\tt 4146cfe2fe9249fc14623f832587efe197ef5d2d refs/stash};
+  \node[text width=0.9\hsize,anchor=north west,yshift=-1mm] at (b3.south west) (b4) {\footnotesize\tt babdda4735ef164b7023be3545860d8b0bae250a HEAD};
+  \node[inner sep=0mm,fit=(b1) (b4),draw,rounded corners,dashed,minimum width=0.9\hsize] (digestbox) {};
+  \node[rotate=-90,draw,rounded corners,fill=white] at (digestbox.east) {\scriptsize Digest};
+\end{tikzpicture}
+\end{center}
+
 \end{document}

### 2017-04-14 article.tex
Trying the \immediate\pdfobj stream trick, but it doesn't seem to go where I want it

--- 2017-04-12 article.tex
@@
 \documentclass{article}
+
+\pdfminorversion=4
+
+\usepackage{fullpage}
+\usepackage{tikz}
+\usetikzlibrary{fit}
+
+\def\DeflateHeader{\begingroup
+  \pdfcompresslevel=0\relax
+  \pdfobjcompresslevel=0\relax
+    \immediate\pdfobj stream
+        file {.DEFLATEDummyHeader}
+\endgroup}
+
+\def\DeflateHeaderObj{\begingroup
+  \pdfcompresslevel=0\relax
+  \pdfobjcompresslevel=0\relax
+    {\catcode`\%=12\immediate\pdfliteral{%RRRRR
+}}\endgroup}
 
@@
 
+The file format for Git bundles doesn't appear to be formally
+specified anywhere, however, inspecting \texttt{bundle.c} reveals that
+it's relatively straightforward:
+\begin{center}
+\begin{tikzpicture}
+  \node[text width=0.8\hsize] at (0,0) (sig) {\texttt{\# v2 git bundle}};
+  \node[inner sep=0mm,fit=(sig),draw,rounded corners,dashed,minimum width=0.8\hsize] (sigbox) {};
+  \node[draw,rounded corners,fill=white] at (sigbox.north) (siglabel) {Git Bundle Signature};
+  \node[text width=0.8\hsize,anchor=north west,yshift=-1mm] at (sig.south west) (b1) {\tt 3aa340a2e3d125ab6703e5c9bdfede2054a9c0c5 refs/heads/master};
+  \node[text width=0.8\hsize,anchor=north west,yshift=-1mm] at (b1.south west) (b2) {\tt 3aa340a2e3d125ab6703e5c9bdfede2054a9c0c5 refs/remotes/origin/master};
+  \node[text width=0.8\hsize,anchor=north west,yshift=-1mm] at (b2.south west) (b3) {\tt 4146cfe2fe9249fc14623f832587efe197ef5d2d refs/stash};
+  \node[text width=0.8\hsize,anchor=north west,yshift=-1mm] at (b3.south west) (b4) {\tt babdda4735ef164b7023be3545860d8b0bae250a HEAD};
+  \node[inner sep=0mm,fit=(b1) (b4),draw,rounded corners,dashed,minimum width=0.8\hsize] (digestbox) {};
+  \node[rotate=-90,draw,rounded corners,fill=white] at (digestbox.east) {Digest};
+\end{tikzpicture}
+\end{center}
+
 \end{document}

### 2017-04-14 fix_oversize_pdf.py
Started with the script for oversize PDFs

@@ -1,0 +1,85 @@
+#!/usr/bin/env python2
+
+import re
+import sys
+
+def parse_obj(pdf_content, logger = None):
+    if logger is None:
+        logger = lambda s : None
+    bytes_skipped = 0
+    for line in pdf_content.splitlines():
+        if line.startswith('%'):
+            bytes_skipped += len(line) + 1
+            continue
+        m = re.match(r"^\s*\d+\s+\d+\s+obj$", line)
+        if not m:
+            bytes_skipped += len(line) + 1
+            continue
+        after_obj = pdf_content[bytes_skipped+len(line)+1:]
+        m = re.match(r"^\s*<<.*?\/Length\s+(\d+)\s*\n.*?>>.*?stream\n", after_obj, re.MULTILINE | re.DOTALL)
+        if m:
+            bytes_up_to_endstream = len(m.group(0)) + int(m.group(1))
+            after_stream = after_obj[bytes_up_to_endstream:]
+            m2 = re.match(r".*?\n*endstream\s*\nendobj\s*\n", after_stream, re.MULTILINE | re.DOTALL)
+            if not m2:
+                logger("Expected an endstream/endobj for \"%s\", but instead got \"%s\"" % (line, after_stream[:10]))
+                exit(1)
+            #print pdf_content[bytes_skipped + len(line) + 1 + bytes_up_to_endstream + len(m2.group(0)):][:50]
+            return bytes_skipped, len(line) + 1 + bytes_up_to_endstream + len(m2.group(0))
+        m2 = re.match(".*?\n?endobj\s*\n", after_obj, re.MULTILINE | re.DOTALL)
+        if m2:
+            return bytes_skipped, len(m2.group(0))
+        else:
+            logger("Error: did not find end of PDF \"%s\"!\n" % line)
+    return None, None
+
+PDF_HEADER = r"%PDF-1.\d\s*\n%\xD0\xD4\xC5\xD8\s*\n"
+
+def fix_pdf(pdf_content, output = None, logger = None):
+    if output is None:
+        output = sys.stdout
+    if logger is None:
+        logger = lambda s : None
+    start_offset = None
+    offset = 0
+    for i in range(len(pdf_content)):
+        m = re.match(PDF_HEADER,pdf_content[i:],re.MULTILINE | re.DOTALL)
+        if m:
+            start_offset = i
+            logger("Found PDF header at offset %d\n" % start_offset)
+            offset = len(m.group(0))
+            break
+    if start_offset is None:
[…]

### 2017-04-14 article.tex
Successful insertion of DEFLATE header PDF objects

@@
 \begin{tikzpicture}
-  \node[text width=0.8\hsize] at (0,0) (sig) {\texttt{\# v2 git bundle}};
+  \node[text width=0.8\hsize] at (0,0) (sig) {\footnotesize\texttt{\# v2 git bundle}};
   \node[inner sep=0mm,fit=(sig),draw,rounded corners,dashed,minimum width=0.8\hsize] (sigbox) {};
-  \node[draw,rounded corners,fill=white] at (sigbox.north) (siglabel) {Git Bundle Signature};
-  \node[text width=0.8\hsize,anchor=north west,yshift=-1mm] at (sig.south west) (b1) {\tt 3aa340a2e3d125ab6703e5c9bdfede2054a9c0c5 refs/heads/master};
+  \node[draw,rounded corners,fill=white] at (sigbox.north) (siglabel) {\footnotesize Git Bundle Signature};
+  \node[text width=0.8\hsize,anchor=north west,yshift=-1mm] at (sig.south west) (b1) {\scriptsize\tt 3aa340a2e3d125ab6703e5c9bdfede2054a9c0c5 refs/heads/master};
   \node[text width=0.8\hsize,anchor=north west,yshift=-1mm] at (b1.south west) (b2) {\tt 3aa340a2e3d125ab6703e5c9bdfede2054a9c0c5 refs/remotes/origin/master};
@@
   \node[inner sep=0mm,fit=(b1) (b4),draw,rounded corners,dashed,minimum width=0.8\hsize] (digestbox) {};
-  \node[rotate=-90,draw,rounded corners,fill=white] at (digestbox.east) {Digest};
+  \node[rotate=-90,draw,rounded corners,fill=white] at (digestbox.east) {\footnotesize Digest};
 \end{tikzpicture}

### 2017-04-14 fix_oversize_pdf.py

@@ -33,6 +33,22 @@ def parse_obj(pdf_content, logger = None):
             logger("Error: did not find end of PDF \"%s\"!\n" % line)
     return None, None
 
+DEFLATE_OBJ_START="%d 0 obj\n<<\n/Length 5\n>>\nstream\n"
+DEFLATE_OBJ_PRE_LEN=len(DEFLATE_OBJ_START) + 2
+DEFLATE_OBJ_END="endstream\nendobj\n"
+
+def calculate_deflate_locations(objects, additional_bytes = 0):
+    if not objects:
+        return []
+    length = additional_bytes
+    for i, obj in enumerate(objects):
+        # should we put a deflate header before object i?
+        start, l = obj
+        if length + l + DEFLATE_OBJ_PRE_LEN > 0xFFFF:
+            return [i] + map(lambda j : j+i, calculate_deflate_locations(objects[i:], additional_bytes = len(DEFLATE_OBJ_END)))
+        length += l
+    return []
+
 PDF_HEADER = r"%PDF-1.\d\s*\n%\xD0\xD4\xC5\xD8\s*\n"
 
 def fix_pdf(pdf_content, output = None, logger = None):
@@ -57,10 +73,23 @@ def fix_pdf(pdf_content, output = None, logger = None):
         if start is None:
             break
         objects.append((offset + start, length))
-        if length > 0xFFFF:
-            raise Exception("The object at PDF offset %d is more than 0xFFFF bytes! This PDF cannot be fixed.")
+        if length > 0xFFFF - DEFLATE_OBJ_PRE_LEN:
+            raise Exception("The object at PDF offset %d is more than %d bytes! This PDF cannot be fixed." % (0xFFFF - DEFLATE_OBJ_PRE_LEN))
         offset += start + length
-    logger("Parsed %d PDF objects." % len(objects))
+    logger("Parsed %d PDF objects.\n" % len(objects))
+    locations = calculate_deflate_locations(objects)
+    if not locations:
+        logger("The PDF doesn't need fixing!\n")
+        return
+    if len(locations) > 10:
+        raise Exception("Error: We currently only support up to 10 DEFLATE header objects! Edit fix_oversize_pdf.py to increase this amount.")
+    for idx, i in enumerate(locations):
+        logger("Inserting a DEFLATE header object before existing object #%d...\n" % (i+1))
[…]

### 2017-04-14 fix_oversize_pdf.py
Added a TODO comment

@@ -89,6 +89,7 @@ def fix_pdf(pdf_content, output = None, logger = None):
         pdf_content = pdf_content[:objects[i][0]] + new_obj + pdf_content[objects[i][0]:]
         for j in range(i,len(objects)):
             objects[j] = (objects[j][0] + len(new_obj), objects[j][1])
+    # TODO: Fix the xrefs!
     output.write(pdf_content)
 
 if __name__ == "__main__":

### 2017-04-14 update_deflate_headers.py
Started another script to update the DEFLATE headers

@@ -1,0 +1,46 @@
+#!/usr/bin/env python2
+
+import fix_oversize_pdf
+import re
+import sys
+
+def read_deflate_header(header):
+    last = bool(0b1 & ord(header[0]))
+    length = (ord(header[2]) << 8) + ord(header[1])
+    nlength = (ord(header[4]) << 8) + ord(header[3])
+    if nlength ^ 0xFFFF != length:
+        raise Exception("Corrupt DEFLATE header!")
+    return last, length
+
+def update_deflate_headers(pdf_content, output):
+    m = re.match(r"(.*?)" + fix_oversize_pdf.PDF_HEADER,pdf_content,re.MULTILINE | re.DOTALL)
+    if not m:
+        raise Exception("Could not find PDF header!")
+    pdf_header_offset = len(m.group(1))
+    print "Found PDF header at offset %d" % pdf_header_offset
+    deflate_header = pdf_content[:pdf_header_offset][-5:]
+    last, length = read_deflate_header(deflate_header)
+    if last:
+        print "The entire PDF fits in a single DEFLATE block; nothing needed!"
+        return
+    print "Deleting the unwanted DEFLATE headers..."
+    header_offset = pdf_header_offset + length
+    while not last:
+        header = pdf_content[header_offset:header_offset+5]
+        try:
+            last, length = read_deflate_header(header)
+        except Exception as e:
+            print " ".join(map(hex, map(ord, pdf_content[header_offset-5:header_offset+10])))
+            raise e
+        pdf_content = pdf_content[:header_offset] + pdf_content[header_offset + 5:]
+        print "Deleted DEFLATE header for a %d byte block at offset 0x%x" % (length, header_offset)
+        header_offset += length
+
+if __name__ == "__main__":
+    if len(sys.argv) != 3:
+        sys.stderr.write("Usage: %s PATH_TO_PDF OUTPUT_FILE\n\n" % sys.argv[0])
+        exit(1)
+
+    with open(sys.argv[1], 'rb') as f:
+        with open(sys.argv[2], 'wb') as out:
+            update_deflate_headers(f.read(), out)

### 2017-04-15 fix_oversize_pdf.py
Changed the DEFLATE object placeholder

@@ -36,6 +36,7 @@ def parse_obj(pdf_content, logger = None):
 DEFLATE_OBJ_START="%d 0 obj\n<<\n/Length 5\n>>\nstream\n"
 DEFLATE_OBJ_PRE_LEN=len(DEFLATE_OBJ_START) + 2
 DEFLATE_OBJ_END="endstream\nendobj\n"
+DEFLATE_OBJ_PLACEHOLDER="\x0E\x0F\x10\x11\x12"
 
 def calculate_deflate_locations(objects, additional_bytes = 0):
     if not objects:
@@ -67,6 +68,9 @@ def fix_pdf(pdf_content, output = None, logger = None):
             break
     if start_offset is None:
         raise Exception("Did not find PDF header!")
+    # Make sure that the DEFLATE object placeholder byte sequence doesn't exist in the PDF anywhere:
+    if DEFLATE_OBJ_PLACEHOLDER in pdf_content[offset:]:
+        raise Exception("Error: The PDF already contains the placeholder DEFLATE byte sequence! To fix this, change DEFLATE_OBJ_PLACEHOLDER in fix_oversize_pdf.py")
     objects = []
     while True:
         start, length = parse_obj(pdf_content[offset:], logger = logger)
@@ -85,7 +89,7 @@ def fix_pdf(pdf_content, output = None, logger = None):
         raise Exception("Error: We currently only support up to 10 DEFLATE header objects! Edit fix_oversize_pdf.py to increase this amount.")
     for idx, i in enumerate(locations):
         logger("Inserting a DEFLATE header object before existing object #%d...\n" % (i+1))
-        new_obj = DEFLATE_OBJ_START % (9000+idx) + "XXXXX" + DEFLATE_OBJ_END
+        new_obj = DEFLATE_OBJ_START % (9000+idx) + DEFLATE_OBJ_PLACEHOLDER + DEFLATE_OBJ_END
         pdf_content = pdf_content[:objects[i][0]] + new_obj + pdf_content[objects[i][0]:]
         for j in range(i,len(objects)):
             objects[j] = (objects[j][0] + len(new_obj), objects[j][1])

### 2017-04-15 update_deflate_headers.py
Integrated the DEFLATE header updater, but it's not proucing correct output yet

@@ -12,6 +12,17 @@ def read_deflate_header(header):
         raise Exception("Corrupt DEFLATE header!")
     return last, length
 
+def make_deflate_header(last, length):
+    header = ["\0"] * 5
+    if last:
+        header[0] = "\x01"
+    header[1] = chr(length & 0xFF)
+    header[2] = chr((length & 0xFF00) >> 8)
+    nlength = length ^ 0xFFFF
+    header[3] = chr(nlength & 0xFF)
+    header[4] = chr((nlength & 0xFF00) >> 8)    
+    return "".join(header)
+
 def update_deflate_headers(pdf_content, output):
     m = re.match(r"(.*?)" + fix_oversize_pdf.PDF_HEADER,pdf_content,re.MULTILINE | re.DOTALL)
     if not m:
@@ -35,6 +46,30 @@ def update_deflate_headers(pdf_content, output):
         pdf_content = pdf_content[:header_offset] + pdf_content[header_offset + 5:]
         print "Deleted DEFLATE header for a %d byte block at offset 0x%x" % (length, header_offset)
         header_offset += length
+    print "Updating the injected DEFLATE headers..."
+    previous_header_offset = pdf_header_offset - 5
+    last = False
+    i = 0
+    while not last:
+        i += 1
+        next_header_offset = pdf_content.find(fix_oversize_pdf.DEFLATE_OBJ_PLACEHOLDER, previous_header_offset+5)
+        if next_header_offset < 0:
+            last = True
+            # Find the end of the PDF:
+            PDF_END_MAGIC = "%%EOF\x0A"
+            next_header_offset = pdf_content.find(PDF_END_MAGIC, previous_header_offset+5)
+            if next_header_offset < 0:
+                raise Exception("Could not find the end of the PDF!")
+            next_header_offset += len(PDF_END_MAGIC)
+        length = next_header_offset - previous_header_offset - 5
+        if length > 0xFFFF:
+            raise Exception("The length of DEFLATE block %d is 0x%x bytes, which is over the maximum of 0xFFFF!" % (i, length))
+        pdf_content = pdf_content[:previous_header_offset] + make_deflate_header(last, length) + pdf_content[previous_header_offset+5:]
+        print "END: %s" % " ".join(map(hex, map(ord, pdf_content[previous_header_offset+length:previous_header_offset+5+length])))
+        previous_header_offset = next_header_offset
[…]

### 2017-04-15 update_deflate_headers.py
Removed an unncessary debug printout

@@ -65,7 +65,6 @@ def update_deflate_headers(pdf_content, output):
         if length > 0xFFFF:
             raise Exception("The length of DEFLATE block %d is 0x%x bytes, which is over the maximum of 0xFFFF!" % (i, length))
         pdf_content = pdf_content[:previous_header_offset] + make_deflate_header(last, length) + pdf_content[previous_header_offset+5:]
-        print "END: %s" % " ".join(map(hex, map(ord, pdf_content[previous_header_offset+length:previous_header_offset+5+length])))
         previous_header_offset = next_header_offset
         print "DEFLATE block header %d%s set to length %d" % (i, [""," (last)"][last], length)
     out.write(pdf_content)

### 2017-04-15 update_deflate_headers.py
Added a sanity check

@@ -34,6 +34,7 @@ def update_deflate_headers(pdf_content, output):
     if last:
         print "The entire PDF fits in a single DEFLATE block; nothing needed!"
         return
+    old_length = length
     print "Deleting the unwanted DEFLATE headers..."
     header_offset = pdf_header_offset + length
     while not last:
@@ -43,6 +44,7 @@ def update_deflate_headers(pdf_content, output):
         except Exception as e:
             print " ".join(map(hex, map(ord, pdf_content[header_offset-5:header_offset+10])))
             raise e
+        old_length += length
         pdf_content = pdf_content[:header_offset] + pdf_content[header_offset + 5:]
         print "Deleted DEFLATE header for a %d byte block at offset 0x%x" % (length, header_offset)
         header_offset += length
@@ -50,6 +52,7 @@ def update_deflate_headers(pdf_content, output):
     previous_header_offset = pdf_header_offset - 5
     last = False
     i = 0
+    new_length = 0
     while not last:
         i += 1
         next_header_offset = pdf_content.find(fix_oversize_pdf.DEFLATE_OBJ_PLACEHOLDER, previous_header_offset+5)
@@ -64,9 +67,12 @@ def update_deflate_headers(pdf_content, output):
         length = next_header_offset - previous_header_offset - 5
         if length > 0xFFFF:
             raise Exception("The length of DEFLATE block %d is 0x%x bytes, which is over the maximum of 0xFFFF!" % (i, length))
+        new_length += length
         pdf_content = pdf_content[:previous_header_offset] + make_deflate_header(last, length) + pdf_content[previous_header_offset+5:]
         previous_header_offset = next_header_offset
         print "DEFLATE block header %d%s set to length %d" % (i, [""," (last)"][last], length)
+    if new_length != old_length:
+        raise Exception("The new length is %d bytes, but the old length was %d. This should never happen!" % (new_length, old_length))
     out.write(pdf_content)
     out.flush()
 

### 2017-04-15 update_deflate_headers.py
Fixed one of the sanity checks

@@ -64,6 +64,9 @@ def update_deflate_headers(pdf_content, output):
             if next_header_offset < 0:
                 raise Exception("Could not find the end of the PDF!")
             next_header_offset += len(PDF_END_MAGIC)
+        else:
+            new_length += 5 # Add the length of the DEFLATE header, since that was included in old_length    
+        assert i == 1 or pdf_content[previous_header_offset:previous_header_offset+5] == fix_oversize_pdf.DEFLATE_OBJ_PLACEHOLDER
         length = next_header_offset - previous_header_offset - 5
         if length > 0xFFFF:
             raise Exception("The length of DEFLATE block %d is 0x%x bytes, which is over the maximum of 0xFFFF!" % (i, length))

### 2017-04-16 fix_oversize_pdf.py
Trying to set the DEFLATE headers before adding to git

@@ -33,6 +33,17 @@ def parse_obj(pdf_content, logger = None):
             logger("Error: did not find end of PDF \"%s\"!\n" % line)
     return None, None
 
+def make_deflate_header(last, length):
+    header = ["\0"] * 5
+    if last:
+        header[0] = "\x01"
+    header[1] = chr(length & 0xFF)
+    header[2] = chr((length & 0xFF00) >> 8)
+    nlength = length ^ 0xFFFF
+    header[3] = chr(nlength & 0xFF)
+    header[4] = chr((nlength & 0xFF00) >> 8)    
+    return "".join(header)
+
 DEFLATE_OBJ_START="%d 0 obj\n<<\n/Length 5\n>>\nstream\n"
 DEFLATE_OBJ_PRE_LEN=len(DEFLATE_OBJ_START) + 2
 DEFLATE_OBJ_END="endstream\nendobj\n"
@@ -68,9 +79,6 @@ def fix_pdf(pdf_content, output = None, logger = None):
             break
     if start_offset is None:
         raise Exception("Did not find PDF header!")
-    # Make sure that the DEFLATE object placeholder byte sequence doesn't exist in the PDF anywhere:
-    if DEFLATE_OBJ_PLACEHOLDER in pdf_content[offset:]:
-        raise Exception("Error: The PDF already contains the placeholder DEFLATE byte sequence! To fix this, change DEFLATE_OBJ_PLACEHOLDER in fix_oversize_pdf.py")
     objects = []
     while True:
         start, length = parse_obj(pdf_content[offset:], logger = logger)
@@ -82,19 +90,29 @@ def fix_pdf(pdf_content, output = None, logger = None):
         offset += start + length
     logger("Parsed %d PDF objects.\n" % len(objects))
     locations = calculate_deflate_locations(objects)
+    first_block_size = 0
     if not locations:
         logger("The PDF doesn't need fixing!\n")
-        return
+        return first_block_size
     if len(locations) > 10:
         raise Exception("Error: We currently only support up to 10 DEFLATE header objects! Edit fix_oversize_pdf.py to increase this amount.")
     for idx, i in enumerate(locations):
-        logger("Inserting a DEFLATE header object before existing object #%d...\n" % (i+1))
-        new_obj = DEFLATE_OBJ_START % (9000+idx) + DEFLATE_OBJ_PLACEHOLDER + DEFLATE_OBJ_END
+        if i >= len(locations) - 1:
+            last = True
+            length = len(pdf_content[objects[i][0]:])
+        else:
[…]

### 2017-04-16 update_deflate_headers.py

@@ -12,18 +12,7 @@ def read_deflate_header(header):
         raise Exception("Corrupt DEFLATE header!")
     return last, length
 
-def make_deflate_header(last, length):
-    header = ["\0"] * 5
-    if last:
-        header[0] = "\x01"
-    header[1] = chr(length & 0xFF)
-    header[2] = chr((length & 0xFF00) >> 8)
-    nlength = length ^ 0xFFFF
-    header[3] = chr(nlength & 0xFF)
-    header[4] = chr((nlength & 0xFF00) >> 8)    
-    return "".join(header)
-
-def update_deflate_headers(pdf_content, output):
+def update_deflate_headers(pdf_content, output, first_block_size):
     m = re.match(r"(.*?)" + fix_oversize_pdf.PDF_HEADER,pdf_content,re.MULTILINE | re.DOTALL)
     if not m:
         raise Exception("Could not find PDF header!")
@@ -34,7 +23,6 @@ def update_deflate_headers(pdf_content, output):
     if last:
         print "The entire PDF fits in a single DEFLATE block; nothing needed!"
         return
-    old_length = length
     print "Deleting the unwanted DEFLATE headers..."
     header_offset = pdf_header_offset + length
     while not last:
@@ -44,46 +32,19 @@ def update_deflate_headers(pdf_content, output):
         except Exception as e:
             print " ".join(map(hex, map(ord, pdf_content[header_offset-5:header_offset+10])))
             raise e
-        old_length += length
         pdf_content = pdf_content[:header_offset] + pdf_content[header_offset + 5:]
         print "Deleted DEFLATE header for a %d byte block at offset 0x%x" % (length, header_offset)
         header_offset += length
-    print "Updating the injected DEFLATE headers..."
-    previous_header_offset = pdf_header_offset - 5
-    last = False
-    i = 0
-    new_length = 0
-    while not last:
-        i += 1
-        next_header_offset = pdf_content.find(fix_oversize_pdf.DEFLATE_OBJ_PLACEHOLDER, previous_header_offset+5)
-        if next_header_offset < 0:
-            last = True
-            # Find the end of the PDF:
-            PDF_END_MAGIC = "%%EOF\x0A"
-            next_header_offset = pdf_content.find(PDF_END_MAGIC, previous_header_offset+5)
-            if next_header_offset < 0:
-                raise Exception("Could not find the end of the PDF!")
[…]

### 2017-04-17 fix_oversize_pdf.py
Trying to put the header in a PDF comment instead of an object stream

@@ -33,21 +33,9 @@ def parse_obj(pdf_content, logger = None):
             logger("Error: did not find end of PDF \"%s\"!\n" % line)
     return None, None
 
-def make_deflate_header(last, length):
-    header = ["\0"] * 5
-    if last:
-        header[0] = "\x01"
-    header[1] = chr(length & 0xFF)
-    header[2] = chr((length & 0xFF00) >> 8)
-    nlength = length ^ 0xFFFF
-    header[3] = chr(nlength & 0xFF)
-    header[4] = chr((nlength & 0xFF00) >> 8)    
-    return "".join(header)
-
 DEFLATE_OBJ_START="%d 0 obj\n<<\n/Length 5\n>>\nstream\n"
 DEFLATE_OBJ_PRE_LEN=len(DEFLATE_OBJ_START) + 2
 DEFLATE_OBJ_END="endstream\nendobj\n"
-DEFLATE_OBJ_PLACEHOLDER="\x0E\x0F\x10\x11\x12"
 
 def calculate_deflate_locations(objects, additional_bytes = 0):
     if not objects:
@@ -106,7 +94,7 @@ def fix_pdf(pdf_content, output = None, logger = None):
         if idx == 0:
             first_block_size = length
         logger("Inserting %s DEFLATE header object for a %d byte block before existing object #%d...\n" % (["a", "the last"][last], length, i+1))
-        new_obj = DEFLATE_OBJ_START % (9000+idx) + make_deflate_header(last, length) + DEFLATE_OBJ_END
+        new_obj = DEFLATE_OBJ_START % (9000+idx) + DEFLATE_OBJ_END
         pdf_content = pdf_content[:objects[i][0]] + new_obj + pdf_content[objects[i][0]:]
         for j in range(i,len(objects)):
             objects[j] = (objects[j][0] + len(new_obj), objects[j][1])
@@ -131,9 +119,9 @@ if __name__ == "__main__":
             if len(sys.argv) > 2:
                 out = open(sys.argv[2], 'wb')
                 kwargs["output"] = out
-            first_block_size = fix_pdf(content, **kwargs)
-            with open(sys.argv[1] + ".first_block_bytes", 'w') as l:
-                l.write("%d" % first_block_size)
+            blocks = fix_pdf(content, **kwargs)
+            with open(sys.argv[1] + ".block_offsets", 'w') as l:
+                l.write(" ".join(blocks))
         finally:
             if out is not None:
                 out.close()

### 2017-04-17 update_deflate_headers.py

@@ -10,7 +10,18 @@ def read_deflate_header(header):
     nlength = (ord(header[4]) << 8) + ord(header[3])
     if nlength ^ 0xFFFF != length:
         raise Exception("Corrupt DEFLATE header!")
-    return last, length
+    return last, Length
+
+def make_deflate_header(last, length):
+    header = ["\0"] * 5
+    if last:
+        header[0] = "\x01"
+    header[1] = chr(length & 0xFF)
+    header[2] = chr((length & 0xFF00) >> 8)
+    nlength = length ^ 0xFFFF
+    header[3] = chr(nlength & 0xFF)
+    header[4] = chr((nlength & 0xFF00) >> 8)    
+    return "".join(header)
 
 def update_deflate_headers(pdf_content, output, first_block_size):
     m = re.match(r"(.*?)" + fix_oversize_pdf.PDF_HEADER,pdf_content,re.MULTILINE | re.DOTALL)

### 2017-04-17 fix_oversize_pdf.py
Comments are now being injected

@@ -51,6 +51,18 @@ def calculate_deflate_locations(objects, additional_bytes = 0):
 
 PDF_HEADER = r"%PDF-1.\d\s*\n%\xD0\xD4\xC5\xD8\s*\n"
 
+def bytes_to_inject(length):
+    """Calculates the number of bytes we need to inject after the DEFLATE header to make its length be valid within the PDF"""
+    if length > 0xFFFF:
+        raise Exception("A PDF object of length %d cannot be fixed!" % length)
+    b1 = length & 0xFF
+    b2 = (length & 0xFF00) >> 8
+    n = ord('\n')
+    if (b1 == n) or (b2 == n) or ((b1 ^ 0xFF) == n) or ((b2 ^ 0xFF) == n):
+        return 1 + bytes_to_inject(length + 1)
+    else:
+        return 0
+
 def fix_pdf(pdf_content, output = None, logger = None):
     if output is None:
         output = sys.stdout
@@ -78,10 +90,11 @@ def fix_pdf(pdf_content, output = None, logger = None):
         offset += start + length
     logger("Parsed %d PDF objects.\n" % len(objects))
     locations = calculate_deflate_locations(objects)
-    first_block_size = 0
+    first_block_size = len(pdf_content[start_offset:])
+    block_offsets = [[-5, first_block_size]]
     if not locations:
         logger("The PDF doesn't need fixing!\n")
-        return first_block_size
+        return block_offsets
     if len(locations) > 10:
         raise Exception("Error: We currently only support up to 10 DEFLATE header objects! Edit fix_oversize_pdf.py to increase this amount.")
     for idx, i in enumerate(locations):
@@ -91,26 +104,32 @@ def fix_pdf(pdf_content, output = None, logger = None):
         else:
             last = False
             length = objects[i+1][0] - objects[i][0]
+        length += 1 # We always have to add a newline to end the comment
+        extra_bytes = bytes_to_inject(length)
+        length += extra_bytes
         if idx == 0:
-            first_block_size = length
-        logger("Inserting %s DEFLATE header object for a %d byte block before existing object #%d...\n" % (["a", "the last"][last], length, i+1))
-        new_obj = DEFLATE_OBJ_START % (9000+idx) + DEFLATE_OBJ_END
+            block_offsets[0][1] = objects[i][0]
[…]

### 2017-04-17 update_deflate_headers.py
PDF comment method is now working, but the git bundle is still corrupt

@@ -10,7 +10,7 @@ def read_deflate_header(header):
     nlength = (ord(header[4]) << 8) + ord(header[3])
     if nlength ^ 0xFFFF != length:
         raise Exception("Corrupt DEFLATE header!")
-    return last, Length
+    return last, length
 
 def make_deflate_header(last, length):
     header = ["\0"] * 5
@@ -23,7 +23,7 @@ def make_deflate_header(last, length):
     header[4] = chr((nlength & 0xFF00) >> 8)    
     return "".join(header)
 
-def update_deflate_headers(pdf_content, output, first_block_size):
+def update_deflate_headers(pdf_content, output, block_offsets):
     m = re.match(r"(.*?)" + fix_oversize_pdf.PDF_HEADER,pdf_content,re.MULTILINE | re.DOTALL)
     if not m:
         raise Exception("Could not find PDF header!")
@@ -45,17 +45,27 @@ def update_deflate_headers(pdf_content, output, first_block_size):
             raise e
         pdf_content = pdf_content[:header_offset] + pdf_content[header_offset + 5:]
         print "Deleted DEFLATE header for a %d byte block at offset 0x%x" % (length, header_offset)
-        header_offset += length
+        header_offset += length 
     print "Updating the first DEFLATE header..."
-    pdf_content = pdf_content[:pdf_header_offset] + fix_oversize_pdf.make_deflate_header(False, first_block_size) + pdf_content[pdf_header_offset+5:]
+    pdf_content = pdf_content[:pdf_header_offset + block_offsets[0][0]] + make_deflate_header(False, block_offsets[0][1]) + pdf_content[pdf_header_offset:]
+    print "Updating the injected DEFLATE headers..."
+    for idx, block in enumerate(block_offsets[1:]):
+        last = (idx == len(block_offsets) - 1)
+        offset, length = block
+        print "Injecting DEFLATE header at offset %d for a block of length %d" % (pdf_header_offset + offset, length)
+        pdf_content = pdf_content[:pdf_header_offset + offset] + make_deflate_header(last, length) + pdf_content[pdf_header_offset + offset:]
     out.write(pdf_content)
     out.flush()
 
 if __name__ == "__main__":
     if len(sys.argv) != 4:
-        sys.stderr.write("Usage: %s PATH_TO_PDF OUTPUT_FILE FIRST_BLOCK_BYTES\n\n" % sys.argv[0])
[…]

### 2017-04-17 fix_oversize_pdf.py
Fixing a bunch of bugs with the DEFLATE fixing

@@ -95,10 +95,8 @@ def fix_pdf(pdf_content, output = None, logger = None):
     if not locations:
         logger("The PDF doesn't need fixing!\n")
         return block_offsets
-    if len(locations) > 10:
-        raise Exception("Error: We currently only support up to 10 DEFLATE header objects! Edit fix_oversize_pdf.py to increase this amount.")
     for idx, i in enumerate(locations):
-        if i >= len(locations) - 1:
+        if idx >= len(locations) - 1:
             last = True
             length = len(pdf_content[objects[i][0]:])
         else:
@@ -108,12 +106,12 @@ def fix_pdf(pdf_content, output = None, logger = None):
         extra_bytes = bytes_to_inject(length)
         length += extra_bytes
         if idx == 0:
-            block_offsets[0][1] = objects[i][0]
-        block_offsets.append([objects[i][0] + 5*idx, length]) # 5*idx is to account for the previously added 5-byte DEFLATE block headers
+            block_offsets[0][1] = objects[i][0] + 3 # +3 for the leading '%% '
+        block_offsets.append([objects[i][0] + 5*idx + 3, length]) # 5*idx is to account for the previously added 5-byte DEFLATE block headers, +3 for the leading '%% '
         logger("Inserting %s DEFLATE header comment for a %d byte block with %d extra bytes before existing object #%d...\n" % (["a", "the last"][last], length - extra_bytes, extra_bytes, i+1))
         new_obj = "%%%% %s\n" % (' ' * extra_bytes)
         pdf_content = pdf_content[:objects[i][0]] + new_obj + pdf_content[objects[i][0]:]
-        for j in range(i+1,len(objects)):
+        for j in range(i,len(objects)):
             objects[j] = (objects[j][0] + len(new_obj), objects[j][1])
     # TODO: Fix the xrefs!
     output.write(pdf_content)

### 2017-04-17 update_deflate_headers.py

@@ -44,15 +44,15 @@ def update_deflate_headers(pdf_content, output, block_offsets):
             print " ".join(map(hex, map(ord, pdf_content[header_offset-5:header_offset+10])))
             raise e
         pdf_content = pdf_content[:header_offset] + pdf_content[header_offset + 5:]
-        print "Deleted DEFLATE header for a %d byte block at offset 0x%x" % (length, header_offset)
+        print "Deleted DEFLATE header at offset 0x%x for a %d byte block" % (header_offset, length)
         header_offset += length 
     print "Updating the first DEFLATE header..."
     pdf_content = pdf_content[:pdf_header_offset + block_offsets[0][0]] + make_deflate_header(False, block_offsets[0][1]) + pdf_content[pdf_header_offset:]
     print "Updating the injected DEFLATE headers..."
     for idx, block in enumerate(block_offsets[1:]):
-        last = (idx == len(block_offsets) - 1)
+        last = (idx == len(block_offsets) - 2)
         offset, length = block
-        print "Injecting DEFLATE header at offset %d for a block of length %d" % (pdf_header_offset + offset, length)
+        print "Injecting DEFLATE header at offset 0x%x for a %d byte block" % (pdf_header_offset + offset, length)
         pdf_content = pdf_content[:pdf_header_offset + offset] + make_deflate_header(last, length) + pdf_content[pdf_header_offset + offset:]
     out.write(pdf_content)
     out.flush()

### 2017-04-17 update_deflate_headers.py
Added a validation step

@@ -3,6 +3,7 @@
 import fix_oversize_pdf
 import re
 import sys
+import zlib
 
 def read_deflate_header(header):
     last = bool(0b1 & ord(header[0]))
@@ -29,6 +30,7 @@ def update_deflate_headers(pdf_content, output, block_offsets):
         raise Exception("Could not find PDF header!")
     pdf_header_offset = len(m.group(1))
     print "Found PDF header at offset %d" % pdf_header_offset
+    content_before = zlib.decompress(pdf_content[pdf_header_offset - 7:])
     deflate_header = pdf_content[:pdf_header_offset][-5:]
     last, length = read_deflate_header(deflate_header)
     if last:
@@ -54,6 +56,10 @@ def update_deflate_headers(pdf_content, output, block_offsets):
         offset, length = block
         print "Injecting DEFLATE header at offset 0x%x for a %d byte block" % (pdf_header_offset + offset, length)
         pdf_content = pdf_content[:pdf_header_offset + offset] + make_deflate_header(last, length) + pdf_content[pdf_header_offset + offset:]
+    content_after = zlib.decompress(pdf_content[pdf_header_offset - 7:])
+    print "Validating the resulting DEFLATE headers..."
+    if content_before != content_after:
+        raise Exception("Error: the updated DEFLATE output is corrupt!")
     out.write(pdf_content)
     out.flush()
 

### 2017-04-17 update_deflate_headers.py
Getting close to correcting the Pack SHA1

@@ -1,10 +1,79 @@
 #!/usr/bin/env python2
 
 import fix_oversize_pdf
+import hashlib
 import re
+import struct
 import sys
 import zlib
 
+def parse_pack_object(data):
+    obj_type = None
+    length = 0
+    header_bytes = 0
+    for c in map(ord, data):
+        if obj_type is None:
+            # This is the first byte
+            obj_type = (c >> 4) & 0b111
+            length = c & 0b1111
+        else:
+            length += (c & 0b01111111) << ((header_bytes - 1)*7 + 4)
+        header_bytes += 1
+        if not (c & 0b10000000):
+            break # This was the last byte
+    if obj_type == 7:
+        # OBJ_REF_DELTA
+        header_bytes += 20
+    elif obj_type == 6:
+        # OBJ_OFS_DELTA
+        for c in map(ord, data[header_bytes:]):
+            header_bytes += 1
+            if not (c & 0b10000000):
+                break
+    d = zlib.decompressobj()
+    decompressed = d.decompress(data[header_bytes:], length)
+    assert len(decompressed) == length
+    compressed_length = len(data) - header_bytes - len(d.unused_data)
+    return header_bytes, obj_type, length, compressed_length
+
+def fix_pack_sha1(pdf_content, pdf_header_offset):
+    pack_offset = pdf_content.rindex("PACK", 0, pdf_header_offset)
+    version, num_objects = struct.unpack("!II", pdf_content[pack_offset + 4:pack_offset + 12])
+    print "Found git pack version %d containing %d objects" % (version, num_objects)
+    offset = pack_offset + 12
+    for i in range(num_objects):
+        #print "Offset: 0x%x" % offset
+        header_bytes, obj_type, decompressed_length, compressed_length = parse_pack_object(pdf_content[offset:])
+        #print "Parsed pack object at offset 0x%x of type %d with a %d byte header, %d byte body (decompressed), and %d byte body (compressed)" % (offset, obj_type, header_bytes, decompressed_length, compressed_length)
+        offset += header_bytes + compressed_length
+    print "SHA1 should be at offset 0x%x" % offset
+    sha1 = hashlib.sha1(pdf_content[:offset])
+    print sha1.hexdigest()
+    if sha1.digest() == pdf_content[offset:offset+20]:
+        print "SHA1 is valid!"
+    else:
+        print "SHA1 is not valid!"
+    return offset
[…]

### 2017-04-17 update_deflate_headers.py
Got the SHA1 repair working!

@@ -34,7 +34,7 @@ def parse_pack_object(data):
     decompressed = d.decompress(data[header_bytes:], length)
     assert len(decompressed) == length
     compressed_length = len(data) - header_bytes - len(d.unused_data)
-    return header_bytes, obj_type, length, compressed_length
+    return header_bytes, obj_type, length, compressed_length, decompressed
 
 def fix_pack_sha1(pdf_content, pdf_header_offset):
     pack_offset = pdf_content.rindex("PACK", 0, pdf_header_offset)
@@ -43,36 +43,18 @@ def fix_pack_sha1(pdf_content, pdf_header_offset):
     offset = pack_offset + 12
     for i in range(num_objects):
         #print "Offset: 0x%x" % offset
-        header_bytes, obj_type, decompressed_length, compressed_length = parse_pack_object(pdf_content[offset:])
+        header_bytes, obj_type, decompressed_length, compressed_length, decompressed = parse_pack_object(pdf_content[offset:])
         #print "Parsed pack object at offset 0x%x of type %d with a %d byte header, %d byte body (decompressed), and %d byte body (compressed)" % (offset, obj_type, header_bytes, decompressed_length, compressed_length)
         offset += header_bytes + compressed_length
     print "SHA1 should be at offset 0x%x" % offset
-    sha1 = hashlib.sha1(pdf_content[:offset])
+    sha1 = hashlib.sha1(pdf_content[pack_offset:offset])
     print sha1.hexdigest()
     if sha1.digest() == pdf_content[offset:offset+20]:
         print "SHA1 is valid!"
     else:
-        print "SHA1 is not valid!"
-    return offset
-    # sha1 = None
-    # last_percent = -1
-    # for i in range(pack_offset + 12 + 20, len(pdf_content) - 20):
-    #     percent = int(float(i - (pack_offset + 12 + 20)) / float(len(pdf_content)-20-(pack_offset + 12 + 20)) * 100.0)
-    #     if percent > last_percent:
-    #         sys.stdout.write('\r' + ' ' * 30 + "\rSearching for end of pack... %d%%" % percent)
-    #         sys.stdout.flush()
-    #         last_percent = percent
-    #     pack_content = pdf_content[pack_offset:i]
-    #     test = hashlib.sha1(pack_content)
-    #     assert len(test.digest()) == 20
-    #     #print test.digest(), pdf_content[i:i+20]
[…]

### 2017-04-27 update_deflate_headers.py
Trying to get pack file reordering working

@@ -27,7 +27,7 @@ def decode_obj_ref(data):
         if not (c & 0b10000000):
             break
     if bytes_read >= 2:
-	reference += (1 << (7 * (bytes_read - 1)))
+        reference += (1 << (7 * (bytes_read - 1)))
     return reference, bytes_read
 
 def encode_obj_ref(offset):
@@ -93,6 +93,36 @@ def parse_pack_object(data):
     kwargs["decompressed"] = decompressed
     return PackObject(**kwargs)
 
+def verify_pack(pdf_content, pdf_header_offset, pdf_size_delta = 0):
+    pack_offset = pdf_content.rindex("PACK", 0, pdf_header_offset)
+    version, num_objects = struct.unpack("!II", pdf_content[pack_offset + 4:pack_offset + 12])
+    print "Found git pack version %d containing %d objects" % (version, num_objects)
+    start_offset = pack_offset + 12
+    offset = start_offset
+    bytes_since_pdf = None
+    pdf_length = None
+    offset_delta = pdf_size_delta
+    objects_by_offset = {}
+    for i in range(num_objects):
+        obj = parse_pack_object(pdf_content[offset:])
+        objects_by_offset[offset] = obj
+        if obj.obj_type == OBJ_OFS_DELTA:
+            print "Checking OBJ OFS DELTA reference..."
+            # Sanity check: make sure that the new reference is correct
+            #if offset - obj.reference not in objects_by_offset:
+            #    try:
+            #        parse_pack_object(pdf_content[offset-obj.reference-pdf_size_delta:])
+            #    except Exception:
+            #        pass
+            #try:
+            #    print offset, obj.reference, pdf_size_delta
+            #    ref = parse_pack_object(pdf_content[offset-obj.reference-pdf_size_delta:])
+            #    print "Valid OBJ_OFS_DELTA pack object, referring to an object of type %s" % ref.obj_type
+            #except zlib.error:
+            #    raise Exception("Delta offset at file offset 0x%x is referencing back %d bytes, but that does not correspond to a Pack object!" % (offset, obj.reference))
+        offset += obj.header_bytes + obj.compressed_length
+    return pack_offset, start_offset
+
 def fix_pack_sha1(pdf_content, pdf_header_offset, fix = False, pdf_size_delta = 0):
     pack_offset = pdf_content.rindex("PACK", 0, pdf_header_offset)
[…]

### 2017-04-28 update_deflate_headers.py
Removed a bunch of code that is now OBE

@@ -93,36 +93,6 @@ def parse_pack_object(data):
     kwargs["decompressed"] = decompressed
     return PackObject(**kwargs)
 
-def verify_pack(pdf_content, pdf_header_offset, pdf_size_delta = 0):
-    pack_offset = pdf_content.rindex("PACK", 0, pdf_header_offset)
-    version, num_objects = struct.unpack("!II", pdf_content[pack_offset + 4:pack_offset + 12])
-    print "Found git pack version %d containing %d objects" % (version, num_objects)
-    start_offset = pack_offset + 12
-    offset = start_offset
-    bytes_since_pdf = None
-    pdf_length = None
-    offset_delta = pdf_size_delta
-    objects_by_offset = {}
-    for i in range(num_objects):
-        obj = parse_pack_object(pdf_content[offset:])
-        objects_by_offset[offset] = obj
-        if obj.obj_type == OBJ_OFS_DELTA:
-            print "Checking OBJ OFS DELTA reference..."
-            # Sanity check: make sure that the new reference is correct
-            #if offset - obj.reference not in objects_by_offset:
-            #    try:
-            #        parse_pack_object(pdf_content[offset-obj.reference-pdf_size_delta:])
-            #    except Exception:
-            #        pass
-            #try:
-            #    print offset, obj.reference, pdf_size_delta
-            #    ref = parse_pack_object(pdf_content[offset-obj.reference-pdf_size_delta:])
-            #    print "Valid OBJ_OFS_DELTA pack object, referring to an object of type %s" % ref.obj_type
-            #except zlib.error:
-            #    raise Exception("Delta offset at file offset 0x%x is referencing back %d bytes, but that does not correspond to a Pack object!" % (offset, obj.reference))
-        offset += obj.header_bytes + obj.compressed_length
-    return pack_offset, start_offset
-
 def fix_pack_sha1(pdf_content, pdf_header_offset, fix = False, pdf_size_delta = 0):
     pack_offset = pdf_content.rindex("PACK", 0, pdf_header_offset)
     version, num_objects = struct.unpack("!II", pdf_content[pack_offset + 4:pack_offset + 12])
@@ -137,34 +107,9 @@ def fix_pack_sha1(pdf_content, pdf_header_offset, fix = False, pdf_size_delta = 
         obj = parse_pack_object(pdf_content[offset:])
[…]

### 2017-04-29 article.tex
Started wiriting the background on Git

--- 2017-04-14 article.tex
@@
 \usepackage{tikz}
-\usetikzlibrary{fit}
+\usetikzlibrary{fit,patterns,decorations.pathreplacing}
+
+\usepackage[T1,T2A]{fontenc}
+\usepackage[utf8]{inputenc}
+\usepackage[russian,english]{babel} %Last language is default and active.
+\usepackage{graphicx}
+
+\def\thindivider{\centerline{\tiny%%
+-- --- --- -- ~~~ --- --- --- ~~~ --- -- --- -- ~~~~~~
+--- --- --- ~~~  -- --- -- ~~~~~~
+--- --- -- ~~~ --- ~~~ -- -- --- -- ~~~ --- --- ---}}
 
@@
 
+\thindivider
+
+Neighbors, it's possible to create a PDF that is also a Git
+repository.
+
+\begin{center}
+\begingroup
+\setbox9=\hbox{\footnotesize\verb|Receiving objects: 100% (174/174), 103.48 KiB | 0 bytes/s, done.|}
+\begin{minipage}{\wd9}
+\footnotesize\begin{verbatim}
+$ git clone article.pdf foo
+Cloning into 'foo'...
+Receiving objects: 100% (174/174), 103.48 KiB | 0 bytes/s, done.
+Resolving deltas: 100% (100/100), done.
+$ cd foo
+$ ls
+article.tex
+article.pdf
+\end{verbatim}
+\end{minipage}
+\endgroup
+\end{center}
+
+\section{The Git Bundle File Format}
+
 The file format for Git bundles doesn't appear to be formally
@@
 it's relatively straightforward:
+\def\returnkey{\hspace*{1pt}{\setbox9=\hbox{$\hookleftarrow$}\tikz\node[text width=\wd9,fill=lightgray,rounded corners=1pt] at (0,0){$\hookleftarrow$};}\hspace*{1pt}}
 \begin{center}
 \begin{tikzpicture}
-  \node[text width=0.9\hsize] at (0,0) (sig) {\footnotesize\texttt{\# v2 git bundle}};
+  \node[text width=0.9\hsize] at (0,0) (sig) {\footnotesize\tt\# v2 git bundle{\tiny\returnkey}};
   \node[inner sep=0mm,fit=(sig),draw,rounded corners,dashed,minimum width=0.9\hsize] (sigbox) {};
-  \node[draw,rounded corners,fill=white] at (sigbox.north) (siglabel) {\scriptsize Git Bundle Signature};
-  \node[text width=0.9\hsize,anchor=north west,yshift=-1mm] at (sig.south west) (b1) {\footnotesize\tt 3aa340a2e3d125ab6703e5c9bdfede2054a9c0c5 refs/heads/master};
-  \node[text width=0.9\hsize,anchor=north west,yshift=-1mm] at (b1.south west) (b2) {\footnotesize\tt 3aa340a2e3d125ab6703e5c9bdfede2054a9c0c5 refs/remotes/origin/master};
-  \node[text width=0.9\hsize,anchor=north west,yshift=-1mm] at (b2.south west) (b3) {\footnotesize\tt 4146cfe2fe9249fc14623f832587efe197ef5d2d refs/stash};
-  \node[text width=0.9\hsize,anchor=north west,yshift=-1mm] at (b3.south west) (b4) {\footnotesize\tt babdda4735ef164b7023be3545860d8b0bae250a HEAD};
+  \node[draw,rounded corners,fill=white] at (sigbox.north) (siglabel) {\footnotesize Git Bundle Signature};
+  \node[text width=0.9\hsize,anchor=north west,yshift=-0.5mm] at (sig.south west) (b1) {\footnotesize\tt 3aa340a2e3d125ab6703e5c9bdfede2054a9c0c5 refs/heads/master{\tiny\returnkey}};
+  \node[text width=0.9\hsize,anchor=north west,yshift=-0.5mm] at (b1.south west) (b2) {\footnotesize\tt 3aa340a2e3d125ab6703e5c9bdfede2054a9c0c5 refs/remotes/origin/master{\tiny\returnkey}};
+  \node[text width=0.9\hsize,anchor=north west,yshift=-0.5mm] at (b2.south west) (b3) {\footnotesize\tt 4146cfe2fe9249fc14623f832587efe197ef5d2d refs/stash{\tiny\returnkey}};
+  \node[text width=0.9\hsize,anchor=north west,yshift=-0.5mm] at (b3.south west) (b4) {\footnotesize\tt babdda4735ef164b7023be3545860d8b0bae250a HEAD{\tiny\returnkey}};
   \node[inner sep=0mm,fit=(b1) (b4),draw,rounded corners,dashed,minimum width=0.9\hsize] (digestbox) {};
-  \node[rotate=-90,draw,rounded corners,fill=white] at (digestbox.east) {\scriptsize Digest};
+  \node[rotate=-90,draw,rounded corners,fill=white] at (digestbox.east) {\footnotesize Digest};
+  \node[text width=0.9\hsize,anchor=north west,yshift=-0.5mm] at (b4.south west) (empty) {{\tiny\returnkey}\footnotesize\ };
+  \node[text width=0.9\hsize,anchor=north west,yshift=-0.5mm] at (empty.south west) (pack) {{\footnotesize\tt PACK}$\ldots$};
+  \node[inner sep=0mm,fit=(pack),draw,rounded corners,dashed,minimum width=0.9\hsize] (packbox) {};
+  \node[draw,rounded corners,fill=white] at (pack.south) {\footnotesize Git Packfile};
 \end{tikzpicture}
+\end{center}
+Git uses its Packfile format to compress the objects in its database,
+as well as to reduce network bandwidth when pushing and pulling, which
+makes it an obvious choice for use inside bundles. This of course
+raises the question: What is the format for a Git Packfile?
+
+Git does have some internal documentation in
+\begin{center}
+  \texttt{Documentation/technical/pack-format.txt}
+\end{center}
+however, it is rather sparse, and does not provide enough detail to
+fully parse the format. The documentation also has some
+``observations'' that suggest it wasn't even written by the file
+format's creator and instead was written by a developer who was later
+trying to make sense of the code.
+
+Luckily, Aditya Mukerjee already had to reverse engineer the file
+format for his GitGo clean-room implementation of Git, and he wrote an
+excellent blog entry about
+it\footnote{\texttt{https://codewords.recurse.com/issues/three/unpacking-git-packfiles}}.
+
+\begin{center}
+{\small
+\def\char#1{\,`\texttt{#1}'\,}
+\setbox9=\hbox{\char{K}}
+\newdimen\bytewidth\bytewidth=\wd9
+\def\byte#1{\hbox to \bytewidth{\hfil\texttt{#1}\hfil}}
+\def\desc#1#2{\hbox to #1\bytewidth{\hfil #2\hfil}}
+\def\underbrace#1{\draw [
+    thick,
+    decoration={
+        brace,
+        mirror,
+        raise=2pt
+    },
+    decorate
+] ([xshift=1pt]#1.base west) -- ([xshift=-1pt]#1.base east) 
+node (#1label) [pos=0.5,anchor=north,yshift=-2pt]}
+\begin{tikzpicture}
+\node[inner sep=0pt,minimum width=4\bytewidth] at (0,0) (pack) {\char{P}\char{A}\char{C}\char{K}};
+\node[inner sep=0pt,anchor=base west] at (pack.base east) (version) {\byte{00}\byte{00}\byte{00}\byte{02}};
+\node[inner sep=0pt,anchor=base west] at (version.base east) (numobj) {\desc{4}{\# objects}};
+\underbrace{pack}{\tiny magic};
+\underbrace{version}{\tiny version};
+\underbrace{numobj}{\tiny big-endian 4 byte int};
+\node[inner sep=0pt,anchor=north west] at (packlabel.south -| pack.west) (chunks) {one data chunk for each object};
+\node[inner sep=0pt,anchor=north west] at ([yshift=-0.5\baselineskip]chunks.south west) (sha1) {20-byte SHA-1 of all the previous data in the pack};
+\end{tikzpicture}}
+\end{center}
+
+I hope you like Soviet holes, dear neighbor, because chasing this
+rabbit might remind you of \foreignlanguage{russian}{Кольская}. If you
+grew bored reading Tristram Shandy, it's safe to skip to the next
+section, 'cause it's turtles all the way down.
+
+\begin{center}
+\includegraphics[width=0.5\hsize]{RazvodityeKrolikov}
+\end{center}
+
+Okay, the next step is to figure out the ``chunk'' format.  The chunk
+header is variable length, and can be as small as one byte. It encodes
+the type of object and the \emph{uncompressed} size of the object. If
+the object is a \textit{delta} (\textit{i.e.}, a diff, as opposed to a
+complete object), the header is followed by either the SHA-1 hash of
+the base object to which the delta should be applied, or a byte
+reference within the packfile for the start of the base object. The
+remainder of the chunk consists of the object data, zlib-compressed.
+\begin{center}
+{\newcount\bitnum\bitnum=0
+\begin{tikzpicture}
+\node[coordinate] at (0,0) (b0) {};
+\foreach \i in {1,0,1,1,0,1,0,0,0,1,0,1,0,1,1,0} {
+  {\newcount\prevbit\prevbit=\bitnum
+  \global\advance\bitnum by 1
+  \node[anchor=west] at (b\the\prevbit.east) (b\the\bitnum) {\i};}
+  \draw ([xshift=1pt]b\the\bitnum.south west) -- ([xshift=-1pt]b\the\bitnum.south east);
+}
+\end{tikzpicture}}
 \end{center}

### 2017-05-01 fix_oversize_pdf.py
Fixed an error printout

@@ -86,7 +86,7 @@ def fix_pdf(pdf_content, output = None, logger = None):
             break
         objects.append((offset + start, length))
         if length > 0xFFFF - DEFLATE_OBJ_PRE_LEN:
-            raise Exception("The object at PDF offset %d is more than %d bytes! This PDF cannot be fixed." % (0xFFFF - DEFLATE_OBJ_PRE_LEN))
+            raise Exception("The object at PDF offset %d is %d bytes, which is more than the maximum of %d! This PDF cannot be fixed." % (offset + start, length, 0xFFFF - DEFLATE_OBJ_PRE_LEN))
         offset += start + length
     logger("Parsed %d PDF objects.\n" % len(objects))
     locations = calculate_deflate_locations(objects)

### 2017-05-01 update_deflate_headers.py

@@ -188,6 +188,8 @@ def update_deflate_headers(pdf_content, output, block_offsets):
         last = (idx == len(block_offsets) - 2)
         offset, length = block
         print "Injecting DEFLATE header at offset 0x%x for a %d byte block" % (pdf_header_offset + offset, length)
+        # Sanity check: the injection position should be preceeded by a PDF comment
+        assert pdf_content[pdf_header_offset + offset - 3:pdf_header_offset + offset] == '%% '
         pdf_content = pdf_content[:pdf_header_offset + offset] + make_deflate_header(last, length) + pdf_content[pdf_header_offset + offset:]
         pdf_size_delta += 5
     content_after = zlib.decompress(pdf_content[pdf_header_offset - 7:])

### 2017-05-01 article.tex
Shrink the Soviet poster to smaller than 64k
Fixed an error printout
Moved the Russian text to a separate PDF to reduce PDF object size

@@
 
-\usepackage[T1,T2A]{fontenc}
-\usepackage[utf8]{inputenc}
-\usepackage[russian,english]{babel} %Last language is default and active.
 \usepackage{graphicx}
@@
 I hope you like Soviet holes, dear neighbor, because chasing this
-rabbit might remind you of \foreignlanguage{russian}{Кольская}. If you
+rabbit might remind you of \includegraphics{kolskaya}. If you
 grew bored reading Tristram Shandy, it's safe to skip to the next
@@
 \begin{center}
-\includegraphics[width=0.5\hsize]{RazvodityeKrolikov}
+%%\includegraphics[width=0.5\hsize]{RazvodityeKrolikov_small}
 \end{center}

### 2017-05-01 update_deflate_headers.py
Added an assertion

@@ -150,8 +150,11 @@ def make_deflate_header(last, length):
     header[2] = chr((length & 0xFF00) >> 8)
     nlength = length ^ 0xFFFF
     header[3] = chr(nlength & 0xFF)
-    header[4] = chr((nlength & 0xFF00) >> 8)    
-    return "".join(header)
+    header[4] = chr((nlength & 0xFF00) >> 8)
+    header = "".join(header)
+    # Sanity check:
+    assert read_deflate_header(header) == (last, length)
+    return header
 
 def update_deflate_headers(pdf_content, output, block_offsets):
     m = re.match(r"(.*?)" + fix_oversize_pdf.PDF_HEADER,pdf_content,re.MULTILINE | re.DOTALL)

### 2017-05-01 fix_oversize_pdf.py
Fixed a bug in calculating DEFLATE block lengths

@@ -44,8 +44,9 @@ def calculate_deflate_locations(objects, additional_bytes = 0):
     for i, obj in enumerate(objects):
         # should we put a deflate header before object i?
         start, l = obj
-        if length + l + DEFLATE_OBJ_PRE_LEN > 0xFFFF:
-            return [i] + map(lambda j : j+i, calculate_deflate_locations(objects[i:], additional_bytes = len(DEFLATE_OBJ_END)))
+        last = (i == len(objects) - 1)
+        if length + l + [len("%% "),0][last] > 0xFFFF/2:
+            return [i] + map(lambda j : j+i, calculate_deflate_locations(objects[i:], additional_bytes = 1))
         length += l
     return []
 
@@ -53,7 +54,7 @@ PDF_HEADER = r"%PDF-1.\d\s*\n%\xD0\xD4\xC5\xD8\s*\n"
 
 def bytes_to_inject(length):
     """Calculates the number of bytes we need to inject after the DEFLATE header to make its length be valid within the PDF"""
-    if length > 0xFFFF:
+    if length > 0xFFFF/2:
         raise Exception("A PDF object of length %d cannot be fixed!" % length)
     b1 = length & 0xFF
     b2 = (length & 0xFF00) >> 8
@@ -85,8 +86,8 @@ def fix_pdf(pdf_content, output = None, logger = None):
         if start is None:
             break
         objects.append((offset + start, length))
-        if length > 0xFFFF - DEFLATE_OBJ_PRE_LEN:
-            raise Exception("The object at PDF offset %d is %d bytes, which is more than the maximum of %d! This PDF cannot be fixed." % (offset + start, length, 0xFFFF - DEFLATE_OBJ_PRE_LEN))
+        if length > 0xFFFF/2 - DEFLATE_OBJ_PRE_LEN:
+            raise Exception("The object at PDF offset %d is %d bytes, which is more than the maximum of %d! This PDF cannot be fixed." % (offset + start, length, 0xFFFF/2 - DEFLATE_OBJ_PRE_LEN))
         offset += start + length
     logger("Parsed %d PDF objects.\n" % len(objects))
     locations = calculate_deflate_locations(objects)
@@ -101,8 +102,11 @@ def fix_pdf(pdf_content, output = None, logger = None):
             length = len(pdf_content[objects[i][0]:])
         else:
             last = False
-            length = objects[i+1][0] - objects[i][0]
+            length = objects[locations[idx+1]][0] - objects[i][0]
[…]

### 2017-05-01 update_deflate_headers.py

@@ -156,6 +156,30 @@ def make_deflate_header(last, length):
     assert read_deflate_header(header) == (last, length)
     return header
 
+def find_deflate_headers(data):
+    last = False
+    offset = 0
+    i = 0
+    while not last:
+        i += 1
+        try:
+            last, length = read_deflate_header(data[offset:offset + 5])
+        except Exception as e:
+            actual_header = None
+            for j in reduce(lambda x, y : x + y, map(lambda n : (-n,n), range(1,0x8000))):
+                try:
+                    last, length = read_deflate_header(data[offset + j:offset + j + 5])
+                    actual_header = j
+                    break
+                except Exception:
+                    pass
+            if actual_header is not None:
+                raise Exception("Bad DEFLATE header (#%d) at offset %d; it looks like the length was off by %d" % (i, offset, actual_header));
+            else:
+                raise Exception("Bad DEFLATE header (#%d) at offset %d" % (i, offset));
+        yield offset, last, length
+        offset += 5 + length
+
 def update_deflate_headers(pdf_content, output, block_offsets):
     m = re.match(r"(.*?)" + fix_oversize_pdf.PDF_HEADER,pdf_content,re.MULTILINE | re.DOTALL)
     if not m:
@@ -166,24 +190,24 @@ def update_deflate_headers(pdf_content, output, block_offsets):
     assert initial_repair == pdf_content # Make sure the input has a valid SHA1
     content_before = zlib.decompress(pdf_content[pdf_header_offset - 7:])
     deflate_header = pdf_content[:pdf_header_offset][-5:]
-    last, length = read_deflate_header(deflate_header)
-    if last:
-        print "The entire PDF fits in a single DEFLATE block; nothing needed!"
-        return
-    print "Deleting the unwanted DEFLATE headers..."
-    header_offset = pdf_header_offset + length
-    pdf_size_delta = 0
-    while not last:
-        header = pdf_content[header_offset:header_offset+5]
-        try:
-            last, length = read_deflate_header(header)
-        except Exception as e:
-            print " ".join(map(hex, map(ord, pdf_content[header_offset-5:header_offset+10])))
-            raise e
+
+    idx = 0
+    for offset, last, length in find_deflate_headers(pdf_content[pdf_header_offset-5:]):
+        if idx == 0:
+            if last:
+                print "The entire PDF fits in a single DEFLATE block; nothing needed!"
+                return
[…]

### 2017-05-01 fix_oversize_pdf.py
It turns out we can have 0xFFFF byte objects, after all!

@@ -45,7 +45,7 @@ def calculate_deflate_locations(objects, additional_bytes = 0):
         # should we put a deflate header before object i?
         start, l = obj
         last = (i == len(objects) - 1)
-        if length + l + [len("%% "),0][last] > 0xFFFF/2:
+        if length + l + [len("%% "),0][last] > 0xFFFF:
             return [i] + map(lambda j : j+i, calculate_deflate_locations(objects[i:], additional_bytes = 1))
         length += l
     return []
@@ -54,7 +54,7 @@ PDF_HEADER = r"%PDF-1.\d\s*\n%\xD0\xD4\xC5\xD8\s*\n"
 
 def bytes_to_inject(length):
     """Calculates the number of bytes we need to inject after the DEFLATE header to make its length be valid within the PDF"""
-    if length > 0xFFFF/2:
+    if length > 0xFFFF:
         raise Exception("A PDF object of length %d cannot be fixed!" % length)
     b1 = length & 0xFF
     b2 = (length & 0xFF00) >> 8
@@ -86,8 +86,8 @@ def fix_pdf(pdf_content, output = None, logger = None):
         if start is None:
             break
         objects.append((offset + start, length))
-        if length > 0xFFFF/2 - DEFLATE_OBJ_PRE_LEN:
-            raise Exception("The object at PDF offset %d is %d bytes, which is more than the maximum of %d! This PDF cannot be fixed." % (offset + start, length, 0xFFFF/2 - DEFLATE_OBJ_PRE_LEN))
+        if length > 0xFFFF - DEFLATE_OBJ_PRE_LEN:
+            raise Exception("The object at PDF offset %d is %d bytes, which is more than the maximum of %d! This PDF cannot be fixed." % (offset + start, length, 0xFFFF - DEFLATE_OBJ_PRE_LEN))
         offset += start + length
     logger("Parsed %d PDF objects.\n" % len(objects))
     locations = calculate_deflate_locations(objects)

### 2017-05-01 fix_oversize_pdf.py
Fixed an off-by-five error

@@ -33,10 +33,6 @@ def parse_obj(pdf_content, logger = None):
             logger("Error: did not find end of PDF \"%s\"!\n" % line)
     return None, None
 
-DEFLATE_OBJ_START="%d 0 obj\n<<\n/Length 5\n>>\nstream\n"
-DEFLATE_OBJ_PRE_LEN=len(DEFLATE_OBJ_START) + 2
-DEFLATE_OBJ_END="endstream\nendobj\n"
-
 def calculate_deflate_locations(objects, additional_bytes = 0):
     if not objects:
         return []
@@ -45,7 +41,7 @@ def calculate_deflate_locations(objects, additional_bytes = 0):
         # should we put a deflate header before object i?
         start, l = obj
         last = (i == len(objects) - 1)
-        if length + l + [len("%% "),0][last] > 0xFFFF:
+        if length + l + [len("%% "),0][last] + 5 > 0xFFFF:
             return [i] + map(lambda j : j+i, calculate_deflate_locations(objects[i:], additional_bytes = 1))
         length += l
     return []
@@ -86,8 +82,8 @@ def fix_pdf(pdf_content, output = None, logger = None):
         if start is None:
             break
         objects.append((offset + start, length))
-        if length > 0xFFFF - DEFLATE_OBJ_PRE_LEN:
-            raise Exception("The object at PDF offset %d is %d bytes, which is more than the maximum of %d! This PDF cannot be fixed." % (offset + start, length, 0xFFFF - DEFLATE_OBJ_PRE_LEN))
+        if length > 0xFFFF - 5 - 4:
+            raise Exception("The object at PDF offset %d is %d bytes, which is more than the maximum of %d! This PDF cannot be fixed." % (offset + start, length, 0xFFFF - 5 - 4))
         offset += start + length
     logger("Parsed %d PDF objects.\n" % len(objects))
     locations = calculate_deflate_locations(objects)

### 2017-05-01 article.tex
It turns out we can have 0xFFFF byte objects, after all!
More text in the article.

@@
 \begin{center}
-%%\includegraphics[width=0.5\hsize]{RazvodityeKrolikov_small}
+\includegraphics[width=0.5\hsize]{RazvodityeKrolikov_small}
 \end{center}
@@
 remainder of the chunk consists of the object data, zlib-compressed.
+
+This is the format of the variable length chunk header:
 \begin{center}
@@
 \node[coordinate] at (0,0) (b0) {};
-\foreach \i in {1,0,1,1,0,1,0,0,0,1,0,1,0,1,1,0} {
+\foreach \i in {1,0,1,1,0,1,0,0,1,1,0,1,0,1,1,0,0,1,0,0,1,0,0,1} {
   {\newcount\prevbit\prevbit=\bitnum
@@
 }
+\draw [
+    thick,
+    decoration={
+        brace,
+        mirror,
+        raise=-2pt
+    },
+    decorate
+] ([xshift=-1pt]b8.north east) -- ([xshift=1pt]b1.north west)
+node [pos=0.5,anchor=south,yshift=-2pt] {\footnotesize first byte};
+\draw [
+    thick,
+    decoration={
+        brace,
+        mirror,
+        raise=-2pt
+    },
+    decorate
+] ([xshift=-1pt]b16.north east) -- ([xshift=1pt]b9.north west)
+node [pos=0.5,anchor=south,yshift=-2pt] {\footnotesize second byte};
+\draw [
+    thick,
+    decoration={
+        brace,
+        mirror,
+        raise=-2pt
+    },
+    decorate
+] ([xshift=-1pt]b24.north east) -- ([xshift=1pt]b17.north west)
+node [pos=0.5,anchor=south,yshift=-2pt] {\footnotesize third byte};
+
+\draw [
+    thick,
+    decoration={
+        brace,
+        mirror,
+        raise=2pt
+    },
+    decorate
+] ([xshift=1pt]b2.south west) -- ([xshift=-1pt]b4.south east)
+node [pos=0.5,yshift=-2pt,anchor=north] (type) {\footnotesize object type};
+
+\draw[<-] (b1) -- (type.south -| b1) node[anchor=north,align=center,font=\footnotesize] {if the MSB is one,\\then this is not\\the last byte};
+
+\draw [
+    thick,
+    decoration={
+        brace,
+        mirror,
+        raise=2pt
+    },
+    decorate
+] ([xshift=1pt]b5.south west) -- ([xshift=-1pt]b8.south east)
+node [pos=0.5,yshift=-2pt,anchor=north,align=center,font=\footnotesize] (length) {first four\\bits of\\the length\\(big-endian)};
+
+\draw[<-] (b9) -- (length.south -| b9) node[anchor=north,align=center,font=\footnotesize] {MSB is one,\\so this is \emph{not} the last byte};
+
+\draw [
+    thick,
+    decoration={
+        brace,
+        mirror,
+        raise=2pt
+    },
+    decorate
+] ([xshift=1pt]b10.south west) -- ([xshift=-1pt]b16.south east)
+node [pos=0.5,yshift=-2pt,anchor=north,align=center,font=\footnotesize] (next) {the next seven\\bits of the length\\(big-endian)};
+
+\draw[<-] (b17) -- (next.south -| b17) node[anchor=north,align=center,font=\footnotesize] {MSB is zero,\\so this is the last byte};
+
+\draw [
+    thick,
+    decoration={
+        brace,
+        mirror,
+        raise=2pt
+    },
+    decorate
+] ([xshift=1pt]b18.south west) -- ([xshift=-1pt]b24.south east)
+node [pos=0.5,yshift=-2pt,anchor=north,align=center,font=\footnotesize] {the next seven\\bits of the length\\(big-endian)};
+
 \end{tikzpicture}}
+\end{center}
+This example header represents an object of type \texttt{0b011}, which
+happens to be a git blob, and an \emph{uncompressed} length of
+$(100_2\,\verb|<<|\,14) +
+(1010110_2\,\verb|<<|\,7) + 1001001_2
+= 76617$ bytes.  Since this is not a delta object, it is immediately
+followed by the zlib-compressed object data.
+
+%% TODO: Talk about the Delta Ref file format
+
+\section{A Minimal Polyglot PoC}
+
+We now know that a git bundle is really just a git packfile with an
+additional header, and a git packfile stores individual objects using
+zlib, which uses the DEFLATE compression algorithm. DEFLATE supports
+zero compression, so if we can store the PDF in a single object (as
+opposed to it being split into deltas), then we could theoretically
+coerce it to be intact within a valid git bundle.
+
+Forcing the PDF into a single object is easy: We just need to add it
+to the repo last, immediately before generating the bundle.
+
+Getting the object to be compressed with zero compression is also
+relatively easy.  That's because git was built in almost religious
+adherence to The UNIX Philosophy: It is architected with hundreds of
+sub commands it calls ``plumbing,'' of which the vast majority you
+will likely have never heard. For example, you might be aware
+that \texttt{git pull} is equivalent to a \texttt{git fetch} followed
+by a \texttt{git merge}. In fact, the \texttt{pull} code is actually
+spawning a new \texttt{git} child process to execute each of those
+subcommands. Likewise, the \texttt{git bundle} command spawns
+a \texttt{git pack-objects} to generate the packfile portion of the
+bundle.  All we need to do is inject the \texttt{--compression=0}
+argument into the list of command line arguments passed
+to \texttt{pack-objects}.  This is a one-line addition
+to \texttt{bundle.c}:\\
+{\footnotesize
+{\color{gray}
+\verb|    argv_array_pushl(&pack_objects.args,|\\
+\verb|                     "pack-objects", "--all-progress-implied",|}\\
+\verb|                     "--compression=0",|\\
+{\color{gray}
+\verb|                     "--stdout", "--thin", "--delta-base-offset",|\\
+\verb|                     NULL);|
+}}
+
+Using our patched version of git, every object stored in the bundle
+will be uncompressed!
+\begin{center}
+\begingroup
+\setbox9=\hbox{\footnotesize\verb|$ git bundle create article_bundle.pdf --all|}
+\begin{minipage}{\wd9}
+\footnotesize\begin{verbatim}
+$ export PATH=/path/to/patched/git:$PATH
+$ git init .
+$ git add article.pdf
+$ git commit article.pdf -m "added"
+$ git bundle create polyglot.pdf --all
+\end{verbatim}
+\end{minipage}
+\endgroup
 \end{center}

### 2017-05-02 article.tex
Removed an unnecessary .
Added some whitespace above the arrows

@@
 
-\draw[<-] (b1) -- (type.south -| b1) node[anchor=north,align=center,font=\footnotesize] {if the MSB is one,\\then this is not\\the last byte};
+\draw[<-] ([yshift=-2pt]b1.south) -- (type.south -| b1) node[anchor=north,align=center,font=\footnotesize] {if the MSB is one,\\then this is not\\the last byte};
 
@@
 
-\draw[<-] (b9) -- (length.south -| b9) node[anchor=north,align=center,font=\footnotesize] {MSB is one,\\so this is \emph{not} the last byte};
+\draw[<-] ([yshift=-2pt]b9.south) -- (length.south -| b9) node[anchor=north,align=center,font=\footnotesize] {MSB is one,\\so this is \emph{not} the last byte};
 
@@
 
-\draw[<-] (b17) -- (next.south -| b17) node[anchor=north,align=center,font=\footnotesize] {MSB is zero,\\so this is the last byte};
+\draw[<-] ([yshift=-2pt]b17.south) -- (next.south -| b17) node[anchor=north,align=center,font=\footnotesize] {MSB is zero,\\so this \emph{is} the last byte};
 
@@
 $ export PATH=/path/to/patched/git:$PATH
-$ git init .
+$ git init
 $ git add article.pdf
@@
 \end{center}
+Any vanilla, un-patched version of git will be able to clone a repo
+from the bundle. It will also be a valid PDF, since virtually all
+PDF readers ignore garbage bytes before and after the PDF.
+
+\section{Generalizing the PoC}
+
+There are, of course, some
 

### 2017-05-02 article.tex
Added a quote

@@
 I hope you like Soviet holes, dear neighbor, because chasing this
-rabbit might remind you of \includegraphics{kolskaya}. If you
-grew bored reading Tristram Shandy, it's safe to skip to the next
-section, 'cause it's turtles all the way down.
+rabbit might remind you of \includegraphics{kolskaya}. I think it is
+useful to describe the git packfile format here, since it is not well
+documented elsewhere. If that doesn't interest you, it's safe to skip
+to the next section.
 
@@
 header is variable length, and can be as small as one byte. It encodes
-the type of object and the \emph{uncompressed} size of the object. If
-the object is a \textit{delta} (\textit{i.e.}, a diff, as opposed to a
-complete object), the header is followed by either the SHA-1 hash of
-the base object to which the delta should be applied, or a byte
-reference within the packfile for the start of the base object. The
-remainder of the chunk consists of the object data, zlib-compressed.
+the object's type and its \emph{uncompressed} size. If the object is
+a \textit{delta} (\textit{i.e.}, a diff, as opposed to a complete
+object), the header is followed by either the SHA-1 hash of the base
+object to which the delta should be applied, or a byte reference
+within the packfile for the start of the base object. The remainder of
+the chunk consists of the object data, zlib-compressed.
 
@@
 \end{center}
-This example header represents an object of type \texttt{0b011}, which
-happens to be a git blob, and an \emph{uncompressed} length of
-$(100_2\,\verb|<<|\,14) +
-(1010110_2\,\verb|<<|\,7) + 1001001_2
-= 76617$ bytes.  Since this is not a delta object, it is immediately
-followed by the zlib-compressed object data.
+The second through fourth most significant bits of the first byte are
+used to store the object type. The remainder of the bytes in the
+header are of the same format as bytes two and three in this
+example. This example header represents an object of type $11_2$,
+which happens to be a git blob, and an \emph{uncompressed} length of
+$(100_2\,\verb|<<|\,14) + (1010110_2\,\verb|<<|\,7) + 1001001_2 =
+76617$ bytes.  Since this is not a delta object, it is immediately
+followed by the zlib-compressed object data. The header does not
+encode the \emph{compressed} size of the object, since the DEFLATE
+encoding can determine the end of the object as it is being
+decompressed.
+
+At this point, if you found The Life and Opinions of Tristram Shandy
+to be boring or frustrating, then it's probably best to skip to the
+next section, 'cause it's turtles all the way down.
+
+{\fontfamily{jkpvos}\selectfont\begin{center}
+\begin{tikzpicture}
+\node at (0,0) (text) {\begin{minipage}{0.8\hsize}\small
+To come at the exact weight of things= in the scientific steel-yard, the fulchrum, [Walter Shandy] would say, should be almost invisible, to avoid all friction from popular tenets=,---without this= the minutiae of philosophy, which should always= turn the balance, will have no weight at all---Knowledge, like matter, he would affirm, was= divisible in infinitum---that the grains= and scruples= were as= much a part of it, as= the gravitation of the whole world.
+\end{minipage}};
+\node[anchor=north east] at (text.north west) {\LARGE``};
+\node[anchor=south west] at (text.south east) {\LARGE''};
+\end{tikzpicture}
+\end{center}}
 

### 2017-05-02 article.tex
More text

@@
 $(100_2\,\verb|<<|\,14) + (1010110_2\,\verb|<<|\,7) + 1001001_2 =
-76617$ bytes.  Since this is not a delta object, it is immediately
+76$,617 bytes.  Since this is not a delta object, it is immediately
 followed by the zlib-compressed object data. The header does not
@@
 that \texttt{git pull} is equivalent to a \texttt{git fetch} followed
-by a \texttt{git merge}. In fact, the \texttt{pull} code is actually
-spawning a new \texttt{git} child process to execute each of those
+by a \texttt{git merge}. In fact, the \texttt{pull} code actually
+spawns a new \texttt{git} child process to execute each of those
 subcommands. Likewise, the \texttt{git bundle} command spawns
-a \texttt{git pack-objects} to generate the packfile portion of the
-bundle.  All we need to do is inject the \texttt{--compression=0}
-argument into the list of command line arguments passed
-to \texttt{pack-objects}.  This is a one-line addition
-to \texttt{bundle.c}:\\
-{\footnotesize
-{\color{gray}
+a \texttt{git pack-objects} child process to generate the packfile
+portion of the bundle.  All we need to do is inject
+the \texttt{--compression=0} argument into the list of command line
+arguments passed to \texttt{pack-objects}.  This is a one-line
+addition to \texttt{bundle.c}:\\ {\footnotesize {\color{gray}
 \verb|    argv_array_pushl(&pack_objects.args,|\\
@@
 
-There are, of course, some
+There are, of course, several limitations to the minimal PoC given in
+the previous section:
+\begin{enumerate}
+
+\item Adobe, being Adobe, will refuse to open the polyglot unless the PDF is version 1.4 or earlier.
+I guess it doesn't like some element of the git bundle signature or digest if it's PDF~1.5?
+Why?
+Because Adobe, that's why.
+
+\item Leaving the entire Git bundle uncompressed is wasteful if the repo contains other files; really, we only need the PDF to be uncompressed.
+
+\item If the PDF is larger than 65,535 bytes---the maximum size of an uncompressed DEFLATE block---then git will inject 5-byte deflate block headers inside the PDF, likely corrupting it.
+
+\item Adobe will also refuse to open the polyglot unless the PDF is near the beginning of the packfile\footnote{Requiring the PDF header to start near the beginning of a file is common for many, but not all, PDF viewers.}.
+
+\end{enumerate}
+
+The first limitation is easy to fix by instructing \LaTeX\ to produce
+a version 1.4~PDF by adding \texttt{\textbackslash pdfminorversion=4}
+to the document.
+
+The second limitation is a simple matter of software engineering,
+adding a command line argument to the \texttt{git bundle} command that
+accepts the hash of the single file to leave uncompressed, and passing
+that hash to \texttt{git pack-objects}. I have created a fork of git
+with this feature, available here:
+\begin{center}
+  \texttt{https://github.com/ESultanik/git/tree/UncompressedPack}
+\end{center}
+
+As an aside, while fixing the second limitation I discovered that if a
+file has multiple PDFs concatenated after one another
+(\textit{i.e.},~a git bundle polyglot with multiple uncompressed PDFs
+in the repo), then the behavior is viewer-dependent: Some viewers will
+render the first PDF, while others will render the last.  That's a fun
+way to generate a PDF that displays completely different content in,
+say, macOS~Preview versus Adobe.
+
+The third limitation is very tricky, and ultimately why this polyglot
+was not used for the PDF of a digital issue of PoC$\|$GTFO. I've a
+solution, but it will not work if the PDF contains any objects
+(\textit{e.g.}, images) that are larger than 65,535 bytes. A universal
+solution would be to break up the image into smaller ones and tile it
+back together, but that is not feasible for a document the size of a
+PoC$\|$GTFO issue.
+
+DEFLATE headers for uncompressed blocks are very simple: The first
+byte encodes whether the following block is the last in the file, the
+next two bytes encode the block length, and the last two bytes are the
+ones' complement of the length.  Therefore, to resolve this issue, all
+we need to do is move all of the DEFLATE headers that zlib created to
+different positions that won't corrupt the PDF, and update their
+lengths accordingly.
 

### 2017-05-02 fix_oversize_pdf.py
Fixed yet another bug in the DEFLATE header location algorithm

@@ -33,17 +33,17 @@ def parse_obj(pdf_content, logger = None):
             logger("Error: did not find end of PDF \"%s\"!\n" % line)
     return None, None
 
-def calculate_deflate_locations(objects, additional_bytes = 0):
+def calculate_deflate_locations(objects, last_location = 0):
     if not objects:
         return []
-    length = additional_bytes
     for i, obj in enumerate(objects):
         # should we put a deflate header before object i?
-        start, l = obj
-        last = (i == len(objects) - 1)
-        if length + l + [len("%% "),0][last] + 5 > 0xFFFF:
-            return [i] + map(lambda j : j+i, calculate_deflate_locations(objects[i:], additional_bytes = 1))
-        length += l
+        if i == len(objects) - 1:
+            next_location = obj[0] + obj[1]
+        else:
+            next_location = objects[i+1][0]
+        if next_location - last_location + 9 > 0xFFFF:
+            return [i] + map(lambda j : j+i, calculate_deflate_locations(objects[i:], last_location = obj[0]))
     return []
 
 PDF_HEADER = r"%PDF-1.\d\s*\n%\xD0\xD4\xC5\xD8\s*\n"

### 2017-05-02 article.tex
Added to the title
More text about DEFLATE

@@
 
-\title{This PDF is a Git Repository\\Containing its Own \LaTeX\ Source}
+\title{This PDF is a Git Repository\\Containing its Own \LaTeX\ Source\\and a Copy of Itself}
 \author{Evan Sultanik}
@@
 \section{The Git Bundle File Format}
+\label{sec:GitBundleFormat}
 
@@
 
-I hope you like Soviet holes, dear neighbor, because chasing this
-rabbit might remind you of \includegraphics{kolskaya}. I think it is
-useful to describe the git packfile format here, since it is not well
-documented elsewhere. If that doesn't interest you, it's safe to skip
-to the next section.
+Although not entirely required to understand the polyglot, I think it
+is useful to describe the git packfile format here, since it is not
+well documented elsewhere. If that doesn't interest you, it's safe to
+skip to the next section.  But if you do proceed, I hope you like
+Soviet holes, dear neighbor, because chasing this rabbit might remind
+you of \includegraphics{kolskaya}.
 
 \begin{center}
-\includegraphics[width=0.5\hsize]{RazvodityeKrolikov_small}
+\includegraphics[width=0.33\hsize]{RazvodityeKrolikov_small}
 \end{center}
@@
 \item Adobe, being Adobe, will refuse to open the polyglot unless the PDF is version 1.4 or earlier.
-I guess it doesn't like some element of the git bundle signature or digest if it's PDF~1.5?
+I guess it doesn't like some element of the git bundle signature or digest if it's PDF~1.5.
 Why?
@@
 
+So, where can we put a 5-byte DEFLATE header such that it won't
+corrupt the PDF? We could use our standard trick of putting it in a
+PDF object stream that we've exploited countless times before to
+enable PoC$\|$GTFO polyglots. The trouble with that is that object
+streams are fixed-length, so once the PDF is decompressed
+(\textit{i.e.},~when a repo is cloned from the git bundle), then all
+of the 5-byte DEFLATE headers will disappear and the object stream
+lengths would all be incorrect. Instead, I chose to use PDF comments,
+which start at any occurrence of the percent sign character~(\%)
+outside a string or stream and continue until the first occurrence of
+a newline.  All of the PDF viewers I tested don't seem to care if
+comments include non-ASCII characters; they seem to simply scan for a
+newline. Therefore, we can inject ``\texttt{\%\textbackslash n}''
+between PDF objects and move the DEFLATE headers there. The only
+caveat is that the DEFLATE header itself can't contain a newline
+byte~(\texttt{0x0A}), otherwise the comment would be ended
+prematurely.  We can resolve that by adding extra spaces to the end of
+the comment, increasing the length of the following DEFLATE block and
+thus increasing the length bytes in the DEFLATE header and avoiding
+the \texttt{0x0A}. The only concession made with this approach is that
+PDF Xref offsets in the deflated version of the PDF will be off by a
+multiple of 5, due to the removed DEFLATE headers. Fortunately, most
+PDF readers can gracefully handle incorrect Xref offsets (at the
+expense of a slower loading time), and this will only affect the PDF
+contained in the repository, \emph{not} the PDF polyglot.
+
 \end{document}

### 2017-05-02 article.tex
More text

@@
 
-So, where can we put a 5-byte DEFLATE header such that it won't
-corrupt the PDF? We could use our standard trick of putting it in a
-PDF object stream that we've exploited countless times before to
-enable PoC$\|$GTFO polyglots. The trouble with that is that object
-streams are fixed-length, so once the PDF is decompressed
-(\textit{i.e.},~when a repo is cloned from the git bundle), then all
-of the 5-byte DEFLATE headers will disappear and the object stream
-lengths would all be incorrect. Instead, I chose to use PDF comments,
-which start at any occurrence of the percent sign character~(\%)
-outside a string or stream and continue until the first occurrence of
-a newline.  All of the PDF viewers I tested don't seem to care if
-comments include non-ASCII characters; they seem to simply scan for a
+Where can we put a 5-byte DEFLATE header such that it won't corrupt
+the PDF? We could use our standard trick of putting it in a PDF object
+stream that we've exploited countless times before to enable
+PoC$\|$GTFO polyglots. The trouble with that is that object streams
+are fixed-length, so once the PDF is decompressed (\textit{i.e.},~when
+a repo is cloned from the git bundle), then all of the 5-byte DEFLATE
+headers will disappear and the object stream lengths would all be
+incorrect. Instead, I chose to use PDF comments, which start at any
+occurrence of the percent sign character~(\%) outside a string or
+stream and continue until the first occurrence of a newline.  All of
+the PDF viewers I tested don't seem to care if comments include
+non-ASCII characters; they seem to simply scan for a
 newline. Therefore, we can inject ``\texttt{\%\textbackslash n}''
@@
 
+As a final step, we need to update the SHA-1 sum at the end of the
+packfile~(\textit{q.v.}~Section~\ref{sec:GitBundleFormat}), since we
+moved the locations of the DEFLATE headers, thus affecting the hash.
+
+At this point, we have all the tools necessary to create a generalized
+PDF/Git Bundle polyglot for \emph{almost} any PDF and git
+repository. The only remaining hurdle is that some viewers require
+that the PDF occur as early in the packfile as possible. At first, I
+considered applying another patch directly the git source code to make
+the uncompressed object first in the packfile. This approach proved to
+be very involved, in part due to git's UNIX design philosophy and
+architecture of generic code reuse. We're already updating the
+packfile's SHA-1 hash due to changing the DEFLATE headers, so instead
+I decided to simply reorder the objects after-the-fact, subsequent to
+the DEFLATE header fix but before we update the hash. The only
+challenge is that moving objects in the packfile has the potential to
+break offset delta objects, since they refer to their base objects via
+a byte offset within the packfile.  Moving the PDF to the beginning
+will break any offset delta objects that occur after the original
+position of the PDF which refer to base objects that occur before the
+original position of the PDF. That's why I had to dive deeper into the
+rabbit hole of the packfile format to understand the delta object
+headers (as you saw at the end of Section~\ref{sec:GitBundleFormat},
+if you were brave enough to finish it), since I originally attempted
+to rewrite the broken ones. Rewriting the broken offset delta objects
+is the \emph{correct} solution, but, in the end, I discovered a much
+simpler way.
+
+\begin{center}
+\begin{tikzpicture}
+\node at (0,0) (text) {\begin{minipage}{0.8\hsize}\small
+As a matter of fact, G-d just questioned my judgement. He said,
+`Terry, are you worthy to be the man who makes the temple? If you are,
+you must answer: Is this [dastardly], or is this divine intellect?'
+\end{minipage}};
+\node[anchor=north east] at (text.north west) {\LARGE``};
+\node[anchor=south west] at (text.south east) {\LARGE''};
+\node[anchor=north east] at (text.south east) (terry) {\setbox9=\hbox{\llap{---}Terry A.\ Davis, creator of TempleOS}\parbox{\wd9}{\raggedright\box9 self-proclaimed ``smartest programmer that's ever lived''}};
+\end{tikzpicture}
+\end{center}
+
+In the previous section, recall that we created the minimal PoC by
+patching the command line arguments to \texttt{pack-objects}.  One of
+the command line arguments that is already passed by default
+is \texttt{--delta-base-offset}. Running \texttt{git help
+pack-objects} reveals the following:
+\begin{center}
+{\footnotesize\setbox9\hbox{\verb|A packed archive can express the base object of a delta as either a|}
+\begin{minipage}{\wd9}
+\begin{verbatim}
+A packed archive can express the base object of a delta as either a
+20-byte object name or as an offset in the stream, but ancient
+versions of Git don't understand the latter. By default, git
+pack-objects only uses the former format for better compatibility.
+This option allows the command to use the latter format for
+compactness. Depending on the average delta chain length, this
+option typically shrinks the resulting packfile by 3-5 per-cent.
+\end{verbatim}
+\end{minipage}}
+\end{center}
+So all we need to do is \emph{remove} the \texttt{--delta-base-offset}
+argument and git will not include any offset delta objects in the
+pack!
+
 \end{document}

### 2017-05-03 article.tex
I wonder how many readers will think I'm British?
Minor wording tweaks

@@
 \end{center}
-Git uses its Packfile format to compress the objects in its database,
-as well as to reduce network bandwidth when pushing and pulling, which
-makes it an obvious choice for use inside bundles. This of course
+Git has another custom format called a \emph{Packfile} that it uses to
+compress the objects in its database, as well as to reduce network
+bandwidth when pushing and pulling. The packfile is therefore an
+obvious choice for storing objects inside bundles. This of course
 raises the question: What is the format for a Git Packfile?
@@
 
-Okay, the next step is to figure out the ``chunk'' format.  The chunk
+Right, the next step is to figure out the ``chunk'' format.  The chunk
 header is variable length, and can be as small as one byte. It encodes
@@
 stream that we've exploited countless times before to enable
-PoC$\|$GTFO polyglots. The trouble with that is that object streams
-are fixed-length, so once the PDF is decompressed (\textit{i.e.},~when
-a repo is cloned from the git bundle), then all of the 5-byte DEFLATE
+PoC$\|$GTFO polyglots. The trouble with that is: Object streams are
+fixed-length, so once the PDF is decompressed (\textit{i.e.},~when a
+repo is cloned from the git bundle), then all of the 5-byte DEFLATE
 headers will disappear and the object stream lengths would all be
@@
 will break any offset delta objects that occur after the original
-position of the PDF which refer to base objects that occur before the
+position of the PDF that refer to base objects that occur before the
 original position of the PDF. That's why I had to dive deeper into the

### 2017-05-03 article.tex
Small tweaks

@@
 \node at (0,0) (text) {\begin{minipage}{0.8\hsize}\small
-To come at the exact weight of things= in the scientific steel-yard, the fulchrum, [Walter Shandy] would say, should be almost invisible, to avoid all friction from popular tenets=,---without this= the minutiae of philosophy, which should always= turn the balance, will have no weight at all---Knowledge, like matter, he would affirm, was= divisible in infinitum---that the grains= and scruples= were as= much a part of it, as= the gravitation of the whole world.
+To come at the exact weight of things= in the scientific steel-yard, the fulchrum, [Walter Shandy] would say, should be almost invisible, to avoid all friction from popular tenets=,---without this= the minuti\ae\ of philosophy, which should always= turn the balance, will have no weight at all---Knowledge, like matter, he would affirm, was= divisible in infinitum---that the grains= and scruples= were as= much a part of it, as= the gravitation of the whole world.
 \end{minipage}};
@@
 byte~(\texttt{0x0A}), otherwise the comment would be ended
-prematurely.  We can resolve that by adding extra spaces to the end of
-the comment, increasing the length of the following DEFLATE block and
-thus increasing the length bytes in the DEFLATE header and avoiding
-the \texttt{0x0A}. The only concession made with this approach is that
-PDF Xref offsets in the deflated version of the PDF will be off by a
-multiple of 5, due to the removed DEFLATE headers. Fortunately, most
-PDF readers can gracefully handle incorrect Xref offsets (at the
-expense of a slower loading time), and this will only affect the PDF
-contained in the repository, \emph{not} the PDF polyglot.
+prematurely.  We can resolve that, if needed, by adding extra spaces
+to the end of the comment, increasing the length of the following
+DEFLATE block and thus increasing the length bytes in the DEFLATE
+header and avoiding the \texttt{0x0A}. The only concession made with
+this approach is that PDF Xref offsets in the deflated version of the
+PDF will be off by a multiple of 5, due to the removed DEFLATE
+headers. Fortunately, most PDF readers can gracefully handle incorrect
+Xref offsets (at the expense of a slower loading time), and this will
+only affect the PDF contained in the repository, \emph{not} the PDF
+polyglot.
 

### 2017-05-05 article.tex
The Temple deserves capitals, no?
Added an in-joke
Added a license.

@@
 As a matter of fact, G-d just questioned my judgement. He said,
-`Terry, are you worthy to be the man who makes the temple? If you are,
+`Terry, are you worthy to be the man who makes The Temple? If you are,
 you must answer: Is this [dastardly], or is this divine intellect?'
@@
 \end{center}
+
+Terry's not the only one who's written a compiler!
 
@@
 
+\section{License}
+
+Copyright \textcopyright\ 2017 Evan A.~Sultanik
+
+Permission is hereby granted, free of charge, to any person obtaining a copy
+of this document and associated source files (the ``Software''), to deal
+in the Software without restriction, including without limitation the rights
+to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+copies of the Software, and to permit persons to whom the Software is
+furnished to do so, subject to the following conditions:
+
+The above copyright notice, this permission notice, and the entire
+contents and history of its associated git repository shall be
+included in all copies or substantial portions of the Software.
+
+\textsc{The Software is provided ``as is'', without warranty of any kind, express or implied, including but not limited to the warranties of merchantability, fitness for a particular purpose and noninfringement.
+In no event shall the authors or copyright holders be liable for any claim, damages or liability, whether in an action of contract, tort or otherwise, arising from, out of or in connection with the Software or the use or other dealings in the software.}
+
 \end{document}

### 2017-05-05 verify_xrefs.py
Added a script to verify xrefs

@@ -1,0 +1,29 @@
+#!/usr/bin/env python2
+
+import os
+import re
+import subprocess
+import tempfile
+
+PDF_HEADER = r"%PDF-1.\d\s*\n%\xD0\xD4\xC5\xD8\s*\n"
+FNULL = open(os.devnull, 'w')
+
+def find_pdf_start(pdf_bytes):
+    m = re.match(r"(.*?)" + PDF_HEADER,pdf_bytes,re.MULTILINE | re.DOTALL)
+    if not m:
+        raise Exception("Could not find PDF header!")
+    return len(m.group(1))
+
+def verify_xrefs(pdf_bytes):
+    pdf_header_offset = find_pdf_start(pdf_bytes)
+    print "Found PDF header at offset %d" % pdf_header_offset
+    with tempfile.NamedTemporaryFile(suffix=".pdf") as tmp:
+        tmp.write(pdf_bytes[pdf_header_offset:])
+        tmp.flush()
+        subprocess.check_call(["/usr/bin/env", "qpdf", "-qdf", tmp.name, '-'], stdout=FNULL)
+        
+if __name__ == "__main__":
+    import sys
+    for path in sys.argv[1:]:
+        with open(path, 'rb') as f:
+            verify_xrefs(f.read())

### 2017-05-07 article.tex
-\relax- instead of -- to prevent an en-dash
Fixed the bottom of Kolskaya from being cut off

@@
 Soviet holes, dear neighbor, because chasing this rabbit might remind
-you of \includegraphics{kolskaya}.
+you of \raisebox{-0.5pt}{\includegraphics{kolskaya}}.
 
@@
 portion of the bundle.  All we need to do is inject
-the \texttt{--compression=0} argument into the list of command line
+the \texttt{-\relax-compression=0} argument into the list of command line
 arguments passed to \texttt{pack-objects}.  This is a one-line

### 2017-05-07 article.tex
Fixed some punctuation

@@
 \node at (0,0) (text) {\begin{minipage}{0.8\hsize}\small
-To come at the exact weight of things= in the scientific steel-yard, the fulchrum, [Walter Shandy] would say, should be almost invisible, to avoid all friction from popular tenets=,---without this= the minuti\ae\ of philosophy, which should always= turn the balance, will have no weight at all---Knowledge, like matter, he would affirm, was= divisible in infinitum---that the grains= and scruples= were as= much a part of it, as= the gravitation of the whole world.
+To come at the exact weight of things= in the scientific steel-yard, the fulchrum, [Walter Shandy] would say, should be almost invisible, to avoid all friction from popular tenets=;---without this= the minuti\ae\ of philosophy, which should always= turn the balance, will have no weight at all.
+Knowledge, like matter, he would affirm, was= divisible in infinitum;---that the grains= and scruples= were as= much a part of it, as= the gravitation of the whole world.
 \end{minipage}};

### 2017-05-07 article.tex
Added an ending

@@
 
+“Thus—thus, my fellow-labourers and associates in this great harvest of our learning, now ripening before our eyes; thus it is, by slow steps of casual increase, that our knowledge physical, metaphysical, physiological, polemical, nautical, mathematical, aenigmatical, technical, biographical, romantical, chemical, and obstetrical, with fifty other branches of it, (most of 'em ending as these do, in ical) have for these two last centuries and more, gradually been creeping upwards towards that Akme of their perfections, from which, if we may form a conjecture from the advances of these last seven years, we cannot possibly be far off.”
+
+Excerpt From: Laurence Sterne. “The Life and Opinions of Tristram Shandy, Gentleman.” iBooks.
+
 %% TODO: Talk about the Delta Ref file format
@@
 
+{\fontfamily{jkpvos}\selectfont\begin{center}
+\begin{minipage}{0.8\hsize}\small
+Thus=---thus=, my fellow-labourers= and associates= in this great harvest of our learning, now ripening before our eyes=; thus it is=, by slow steps= of casual increase, that our knowledge physical, metaphysical, physiological, polemical, nautical, mathematical, \ae{}nigmatical, technical, biographical, romantical, chemical, obstetrical, and polyglottical, with fifty other branches= of it, (most of 'em ending as= these do, in ical) have for these four last centuries= and more, gradually been creeping upwards= towards that Akme of their perfections=, from which, if we may form a conjecture from the advances= of these last \thepage~pages=, we cannot possibly be far off.
+\end{minipage}
+\end{center}}
+
 \section{License}

### 2017-05-07 article.tex
labourers --> neighbours

--- 2017-05-07 article.tex
@@
 
+{\fontfamily{jkpvos}\selectfont\begin{center}
+\begin{minipage}{0.8\hsize}\small
+Thus=---thus=, my fellow-neighbours= and associates= in this great harvest of our learning, now ripening before our eyes=; thus it is=, by slow steps= of casual increase, that our knowledge physical, metaphysical, physiological, polemical, nautical, mathematical, \ae{}nigmatical, technical, biographical, romantical, chemical, obstetrical, and polyglottical, with fifty other branches= of it, (most of 'em ending as= these do, in ical) have for these four last centuries= and more, gradually been creeping upwards= towards that Akme of their perfections=, from which, if we may form a conjecture from the advances= of these last \thepage~pages=, we cannot possibly be far off.
+\end{minipage}
+\end{center}}
+
 \section{License}

### 2017-05-08 article.tex
Missed a few final `s'es

--- 2017-05-07 article.tex
@@
 \begin{minipage}{0.8\hsize}\small
-Thus=---thus=, my fellow-labourers= and associates= in this great harvest of our learning, now ripening before our eyes=; thus it is=, by slow steps= of casual increase, that our knowledge physical, metaphysical, physiological, polemical, nautical, mathematical, \ae{}nigmatical, technical, biographical, romantical, chemical, obstetrical, and polyglottical, with fifty other branches= of it, (most of 'em ending as= these do, in ical) have for these four last centuries= and more, gradually been creeping upwards= towards that Akme of their perfections=, from which, if we may form a conjecture from the advances= of these last \thepage~pages=, we cannot possibly be far off.
+Thus=---thus=, my fellow-neighbours= and associates= in this= great harvest of our learning, now ripening before our eyes=; thus= it is=, by slow steps= of casual increase, that our knowledge physical, metaphysical, physiological, polemical, nautical, mathematical, \ae{}nigmatical, technical, biographical, romantical, chemical, obstetrical, and polyglottical, with fifty other branches= of it, (most of 'em ending as= these do, in ical) have for these four last centuries= and more, gradually been creeping upwards= towards= that Akme of their perfections=, from which, if we may form a conjecture from the advances= of these last \thepage~pages=, we cannot possibly be far off.
 \end{minipage}

### 2017-05-08 article.tex
Fixed an off-by-one error.

--- 2017-05-07 article.tex
@@
 \begin{minipage}{0.8\hsize}\small
-Thus=---thus=, my fellow-labourers= and associates= in this great harvest of our learning, now ripening before our eyes=; thus it is=, by slow steps= of casual increase, that our knowledge physical, metaphysical, physiological, polemical, nautical, mathematical, \ae{}nigmatical, technical, biographical, romantical, chemical, obstetrical, and polyglottical, with fifty other branches= of it, (most of 'em ending as= these do, in ical) have for these four last centuries= and more, gradually been creeping upwards= towards that Akme of their perfections=, from which, if we may form a conjecture from the advances= of these last \thepage~pages=, we cannot possibly be far off.
+Thus=---thus=, my fellow-neighbours= and associates= in this= great harvest of our learning, now ripening before our eyes=; thus= it is=, by slow steps= of casual increase, that our knowledge physical, metaphysical, physiological, polemical, nautical, mathematical, \ae{}nigmatical, technical, biographical, romantical, chemical, obstetrical, and polyglottical, with fifty other branches= of it, (most of 'em ending as= these do, in ical) have for these four last centuries= and more, gradually been creeping upwards= towards= that Akme of their perfections=, from which, if we may form a conjecture from the advances= of these last \pgfmathparse{int(\thepage + 1)}\pgfmathresult~pages=, we cannot possibly be far off.
 \end{minipage}

### 2017-05-08 article.tex
Filled in the section on Delta Objects

@@
 
-%% TODO: Talk about the Delta Ref file format
+There are two types of delta objects: \emph{references}~(object
+type~7) and \emph{offsets}~(object type~6).  Reference delta objects
+contain an additional 20~bytes at the end of the header before the
+zlib-compressed delta data. These 20~bytes contain the SHA-1 hash of
+the base object to which the delta should be applied.  Offset delta
+objects are exactly the same, however, instead of referencing the base
+object by its SHA-1 hash, it is instead represented by a negative byte
+offset to the start of the object within the pack file.  Since a
+negative byte offset can typically be encoded in two or three bytes,
+it's significantly smaller than a 20-byte SHA-1 hash.  One must
+understand how these offset delta objects are encoded if---say, for
+some strange, masochistic reason---one wanted to change the order of
+objects within a packfile, since doing so would break the negative
+offsets. (Foreshadowing!)
+
+One would
+\emph{think} that git would use the same multi-byte length encoding that they
+used for the uncompressed object length.  But no! This is what we have
+to go off of from the git documentation:
+\begin{center}
+\begingroup
+\setbox9=\hbox{\footnotesize\verb|for n >= 2 adding 2^7 + 2^14 + ... + 2^(7*(n-1))|}
+\begin{minipage}{\wd9}
+\footnotesize\begin{verbatim}
+n bytes with MSB set in all but the last one.
+The offset is then the number constructed by
+concatenating the lower 7 bit of each byte, and
+for n >= 2 adding 2^7 + 2^14 + ... + 2^(7*(n-1))
+to the result.
+\end{verbatim}
+\end{minipage}
+\endgroup
+\end{center}
+Right. Some experimenting resulted in the following decoding logic
+that appears to work:
+\begin{center}
+\begingroup
+\setbox9=\hbox{\footnotesize\verb|        reference += (1 << (7 * (bytes_read - 1)))|}
+\begin{minipage}{\wd9}
+\footnotesize
+{\color{blue}\verb|def|}\verb| decode_obj_ref|{\color{gray}\verb|(|}\verb|data|{\color{gray}\verb|):|}\\
+\verb|    bytes_read = 0|\\
+\verb|    reference = 0|\\
+\verb|    |{\color{blue}\verb|for|}\verb| c |{\color{blue}\verb|in map|}{\color{gray}\verb|(|}\verb|ord, data|{\color{gray}\verb|):|}\\
+\verb|        bytes_read += 1|\\
+\verb|        reference <<= 7|\\
+\verb|        reference += c & 0b01111111|\\
+\verb|        |{\color{blue}\verb|if not|}\verb| |{\color{gray}\verb|(|}\verb|c & 0b10000000|{\color{gray}\verb|):|}\\
+{\color{blue}\verb|            break|}\\
+{\color{blue}\verb|    if|}\verb| bytes_read >= 2|{\color{gray}\verb|:|}\\
+\verb|        reference += |{\color{gray}\verb|(|}\verb|1 << |{\color{gray}\verb|(|}\verb|7 * |{\color{gray}\verb|(|}\verb|bytes_read - 1|{\color{gray}\verb|)))|}\\
+\verb|    return reference, bytes_read|\\
+\end{minipage}
+\endgroup
+\end{center}
+
+The rabbit hole is deeper still; we haven't yet discovered the content
+of the compressed delta objects, let alone how they are applied to
+base objects.  At this point, we have more than sufficient knowledge
+to proceed with the PoC, and my canary died ages ago. Aditya Mukerjee
+did a good job of explaining the process of applying deltas in his
+blog post, so I will stop here and proceed with the polyglot.
 

### 2017-05-09 fix_oversize_pdf.py
First cut at fixing the xrefs

@@ -88,7 +88,7 @@ def fix_pdf(pdf_content, output = None, logger = None):
     logger("Parsed %d PDF objects.\n" % len(objects))
     locations = calculate_deflate_locations(objects)
     first_block_size = len(pdf_content[start_offset:])
-    block_offsets = [[-5, first_block_size]]
+    block_offsets = [[-5, first_block_size, 0]]
     if not locations:
         logger("The PDF doesn't need fixing!\n")
         return block_offsets
@@ -99,22 +99,48 @@ def fix_pdf(pdf_content, output = None, logger = None):
         else:
             last = False
             length = objects[locations[idx+1]][0] - objects[i][0]
-        length += 1 # We always have to add a newline to end the comment
+        extra_bytes = 1 # We always have to add a newline to end the comment
         if idx < len(locations) - 1:
             # We need to add three more bytes for the '%% ' before the next deflate block header:
-            length += 3
-        extra_bytes = bytes_to_inject(length)
+            extra_bytes += 3
+        extra_bytes += bytes_to_inject(length + extra_bytes)
         length += extra_bytes
         if idx == 0:
             block_offsets[0][1] = objects[i][0] + 3 # +3 for the leading '%% '
-        block_offsets.append([objects[i][0] + 5*idx + 3, length]) # 5*idx is to account for the previously added 5-byte DEFLATE block headers, +3 for the leading '%% '
+        block_offsets.append([objects[i][0] + 5*idx + 3, length, extra_bytes]) # 5*idx is to account for the previously added 5-byte DEFLATE block headers, +3 for the leading '%% '
         logger("Inserting %s DEFLATE header comment for a %d byte block with %d extra bytes before existing object #%d...\n" % (["a", "the last"][last], length - extra_bytes, extra_bytes, i+1))
         new_obj = "%%%% %s\n" % (' ' * extra_bytes)
         pdf_content = pdf_content[:objects[i][0]] + new_obj + pdf_content[objects[i][0]:]
         for j in range(i,len(objects)):
             objects[j] = (objects[j][0] + len(new_obj), objects[j][1])
-    # TODO: Fix the xrefs!
-    output.write(pdf_content)
+    # Fix the xrefs:
+    pdf = pdf_content.splitlines(True)
+    xrefoff = pdf.index("xref\n")
[…]

### 2017-05-09 update_deflate_headers.py

@@ -213,7 +213,7 @@ def update_deflate_headers(pdf_content, output, block_offsets):
     print "Updating the injected DEFLATE headers..."
     for idx, block in enumerate(block_offsets[1:]):
         last = (idx == len(block_offsets) - 2)
-        offset, length = block
+        offset, length, added_bytes = block
         print "Injecting DEFLATE header at offset 0x%x for a %d byte block" % (pdf_header_offset + offset, length)
         # Sanity check: the injection position should be preceeded by a PDF comment
         assert pdf_content[pdf_header_offset + offset - 3:pdf_header_offset + offset] == '%% '

### 2017-05-09 fix_oversize_pdf.py
Fixed a byte calculation bug

@@ -99,15 +99,16 @@ def fix_pdf(pdf_content, output = None, logger = None):
         else:
             last = False
             length = objects[locations[idx+1]][0] - objects[i][0]
-        extra_bytes = 1 # We always have to add a newline to end the comment
+        injected_length = 1 # We always have to add a newline to end the comment
         if idx < len(locations) - 1:
             # We need to add three more bytes for the '%% ' before the next deflate block header:
-            extra_bytes += 3
-        extra_bytes += bytes_to_inject(length + extra_bytes)
+            injected_length += 3
+        extra_bytes = bytes_to_inject(length + injected_length)
+        injected_length += extra_bytes
         length += extra_bytes
         if idx == 0:
             block_offsets[0][1] = objects[i][0] + 3 # +3 for the leading '%% '
-        block_offsets.append([objects[i][0] + 5*idx + 3, length, extra_bytes]) # 5*idx is to account for the previously added 5-byte DEFLATE block headers, +3 for the leading '%% '
+        block_offsets.append([objects[i][0] + 5*idx + 3, length, injected_length]) # 5*idx is to account for the previously added 5-byte DEFLATE block headers, +3 for the leading '%% '
         logger("Inserting %s DEFLATE header comment for a %d byte block with %d extra bytes before existing object #%d...\n" % (["a", "the last"][last], length - extra_bytes, extra_bytes, i+1))
         new_obj = "%%%% %s\n" % (' ' * extra_bytes)
         pdf_content = pdf_content[:objects[i][0]] + new_obj + pdf_content[objects[i][0]:]

### 2017-05-09 fix_oversize_pdf.py
Fixed yet another byte length error

@@ -105,7 +105,7 @@ def fix_pdf(pdf_content, output = None, logger = None):
             injected_length += 3
         extra_bytes = bytes_to_inject(length + injected_length)
         injected_length += extra_bytes
-        length += extra_bytes
+        length += injected_length
         if idx == 0:
             block_offsets[0][1] = objects[i][0] + 3 # +3 for the leading '%% '
         block_offsets.append([objects[i][0] + 5*idx + 3, length, injected_length]) # 5*idx is to account for the previously added 5-byte DEFLATE block headers, +3 for the leading '%% '
@@ -135,7 +135,7 @@ def fix_pdf(pdf_content, output = None, logger = None):
         ref[0] = fixed_offset
         output.write(" ".join(ref))
         pdf.pop(0)
-    print "%d bytes added to the last block due to the xref fixes" % xref_len_diff
+    #print "%d bytes added to the last block due to the xref fixes" % xref_len_diff
     block_offsets[-1][1] += xref_len_diff
     xrefoff = pdf.index("startxref\n")
     output.write("".join(pdf[:xrefoff+1]))

### 2017-05-09 fix_oversize_pdf.py
Forgot to uncomment the line that actually updates the xrefs!

@@ -128,7 +128,7 @@ def fix_pdf(pdf_content, output = None, logger = None):
         ref = pdf[0].split(" ")
         idx = int(ref[0])
         while block_idx < len(block_offsets) and block_offsets[block_idx][0] - offset <= idx:
-            #offset += block_offsets[block_idx-1][2]
+            offset += block_offsets[block_idx-1][2]
             block_idx += 1
         fixed_offset = "%010i" % (idx + offset)
         xref_len_diff += len(fixed_offset) - len(ref[0])

### 2017-05-09 fix_oversize_pdf.py
Forgot to account for the PDF header offset

@@ -127,7 +127,7 @@ def fix_pdf(pdf_content, output = None, logger = None):
     for i in range(nxref-1):
         ref = pdf[0].split(" ")
         idx = int(ref[0])
-        while block_idx < len(block_offsets) and block_offsets[block_idx][0] - offset <= idx:
+        while block_idx < len(block_offsets) and block_offsets[block_idx][0] - start_offset - offset <= idx:
             offset += block_offsets[block_idx-1][2]
             block_idx += 1
         fixed_offset = "%010i" % (idx + offset)

### 2017-05-09 fix_oversize_pdf.py
Off-by-one error. There's always an off-by-one error!

@@ -128,7 +128,7 @@ def fix_pdf(pdf_content, output = None, logger = None):
         ref = pdf[0].split(" ")
         idx = int(ref[0])
         while block_idx < len(block_offsets) and block_offsets[block_idx][0] - start_offset - offset <= idx:
-            offset += block_offsets[block_idx-1][2]
+            offset += block_offsets[block_idx][2]
             block_idx += 1
         fixed_offset = "%010i" % (idx + offset)
         xref_len_diff += len(fixed_offset) - len(ref[0])

### 2017-05-09 fix_oversize_pdf.py
Skip the first block, since it doesn't add any extra bytes!

@@ -122,7 +122,7 @@ def fix_pdf(pdf_content, output = None, logger = None):
     output.write("".join(pdf[:xrefoff+3]))
     pdf = pdf[xrefoff+3:]
     offset = 0
-    block_idx = 0
+    block_idx = 1
     xref_len_diff = 0
     for i in range(nxref-1):
         ref = pdf[0].split(" ")

### 2017-05-09 fix_oversize_pdf.py
I was incorrectly assuming that xrefs are sorted according to offset!

@@ -92,6 +92,7 @@ def fix_pdf(pdf_content, output = None, logger = None):
     if not locations:
         logger("The PDF doesn't need fixing!\n")
         return block_offsets
+    original_block_offsets = [-5] + map(lambda i : objects[i][0], locations)
     for idx, i in enumerate(locations):
         if idx >= len(locations) - 1:
             last = True
@@ -99,16 +100,15 @@ def fix_pdf(pdf_content, output = None, logger = None):
         else:
             last = False
             length = objects[locations[idx+1]][0] - objects[i][0]
-        injected_length = 1 # We always have to add a newline to end the comment
+        length += 1 # We always have to add a newline to end the comment
         if idx < len(locations) - 1:
             # We need to add three more bytes for the '%% ' before the next deflate block header:
-            injected_length += 3
-        extra_bytes = bytes_to_inject(length + injected_length)
-        injected_length += extra_bytes
-        length += injected_length
+            length += 3
+        extra_bytes = bytes_to_inject(length)
+        length += extra_bytes
         if idx == 0:
             block_offsets[0][1] = objects[i][0] + 3 # +3 for the leading '%% '
-        block_offsets.append([objects[i][0] + 5*idx + 3, length, injected_length]) # 5*idx is to account for the previously added 5-byte DEFLATE block headers, +3 for the leading '%% '
+        block_offsets.append([objects[i][0] + 5*idx + 3, length, 9 + extra_bytes]) # 5*idx is to account for the previously added 5-byte DEFLATE block headers, +3 for the leading '%% '
         logger("Inserting %s DEFLATE header comment for a %d byte block with %d extra bytes before existing object #%d...\n" % (["a", "the last"][last], length - extra_bytes, extra_bytes, i+1))
         new_obj = "%%%% %s\n" % (' ' * extra_bytes)
         pdf_content = pdf_content[:objects[i][0]] + new_obj + pdf_content[objects[i][0]:]
@@ -121,15 +121,16 @@ def fix_pdf(pdf_content, output = None, logger = None):
     nxref = int(pdf[xrefoff+1].split(" ")[1])
     output.write("".join(pdf[:xrefoff+3]))
     pdf = pdf[xrefoff+3:]
-    offset = 0
-    block_idx = 1
     xref_len_diff = 0
     for i in range(nxref-1):
         ref = pdf[0].split(" ")
         idx = int(ref[0])
[…]

### 2017-05-09 fix_oversize_pdf.py
Account for the start offset of the PDF

@@ -127,7 +127,7 @@ def fix_pdf(pdf_content, output = None, logger = None):
         idx = int(ref[0])
         offset = 0
         block_idx = 1
-        while block_idx < len(block_offsets) and original_block_offsets[block_idx] <= idx:
+        while block_idx < len(block_offsets) and original_block_offsets[block_idx] - start_offset <= idx:
             offset += block_offsets[block_idx][2]
             block_idx += 1
         sys.stdout.write("Increasing xref %d by %d bytes to %d\n" % (i, offset, idx + offset))

### 2017-05-09 fix_oversize_pdf.py
Added some sanity checks

@@ -76,6 +76,7 @@ def fix_pdf(pdf_content, output = None, logger = None):
             break
     if start_offset is None:
         raise Exception("Did not find PDF header!")
+    original_pdf = pdf_content
     objects = []
     while True:
         start, length = parse_obj(pdf_content[offset:], logger = logger)
@@ -123,6 +124,7 @@ def fix_pdf(pdf_content, output = None, logger = None):
     pdf = pdf[xrefoff+3:]
     xref_len_diff = 0
     for i in range(nxref-1):
+        assert len(pdf[0]) == 20 # xref entries must be exactly 20 bytes long, including newlines
         ref = pdf[0].split(" ")
         idx = int(ref[0])
         offset = 0
@@ -132,9 +134,12 @@ def fix_pdf(pdf_content, output = None, logger = None):
             block_idx += 1
         sys.stdout.write("Increasing xref %d by %d bytes to %d\n" % (i, offset, idx + offset))
         fixed_offset = "%010i" % (idx + offset)
+        #assert pdf_content[start_offset + idx + offset/9*4:start_offset + idx + offset/9*4+20] == original_pdf[start_offset + idx:start_offset + idx + 20]
         xref_len_diff += len(fixed_offset) - len(ref[0])
         ref[0] = fixed_offset
-        output.write(" ".join(ref))
+        new_xref = " ".join(ref)
+        assert len(new_xref) == 20
+        output.write(new_xref)
         pdf.pop(0)
     #print "%d bytes added to the last block due to the xref fixes" % xref_len_diff
     block_offsets[-1][1] += xref_len_diff

### 2017-05-09 fix_oversize_pdf.py
Added another fix, but it shouldn't ever be needed

@@ -145,7 +145,7 @@ def fix_pdf(pdf_content, output = None, logger = None):
     block_offsets[-1][1] += xref_len_diff
     xrefoff = pdf.index("startxref\n")
     output.write("".join(pdf[:xrefoff+1]))
-    output.write("%i\n" % (int(pdf[xrefoff+1]) + offset))
+    output.write("%i\n" % (int(pdf[xrefoff+1]) + offset + xref_len_diff))
     pdf=pdf[xrefoff+2:]
     output.write("".join(pdf))
     return block_offsets

### 2017-05-10 verify_xrefs.py
Added a success printout

@@ -21,6 +21,8 @@ def verify_xrefs(pdf_bytes):
         tmp.write(pdf_bytes[pdf_header_offset:])
         tmp.flush()
         subprocess.check_call(["/usr/bin/env", "qpdf", "-qdf", tmp.name, '-'], stdout=FNULL)
+    print "PDF appears to be valid."
+    return True
         
 if __name__ == "__main__":
     import sys

### 2017-05-10 verify_xrefs.py
Ignore junk after %%EOF

@@ -6,19 +6,25 @@ import subprocess
 import tempfile
 
 PDF_HEADER = r"%PDF-1.\d\s*\n%\xD0\xD4\xC5\xD8\s*\n"
+EOF_STRING = '\n%%EOF\n'
 FNULL = open(os.devnull, 'w')
 
 def find_pdf_start(pdf_bytes):
     m = re.match(r"(.*?)" + PDF_HEADER,pdf_bytes,re.MULTILINE | re.DOTALL)
     if not m:
-        raise Exception("Could not find PDF header!")
+        raise Exception('Could not find PDF header!')
     return len(m.group(1))
+
+def find_pdf_end(pdf_bytes, start_offset = 0):
+    return pdf_bytes.rindex(EOF_STRING, start_offset) + len(EOF_STRING)
 
 def verify_xrefs(pdf_bytes):
     pdf_header_offset = find_pdf_start(pdf_bytes)
     print "Found PDF header at offset %d" % pdf_header_offset
+    pdf_end_offset = find_pdf_end(pdf_bytes, start_offset = pdf_header_offset)
+    print "Found PDF trailer at offset %d" % pdf_end_offset
     with tempfile.NamedTemporaryFile(suffix=".pdf") as tmp:
-        tmp.write(pdf_bytes[pdf_header_offset:])
+        tmp.write(pdf_bytes[pdf_header_offset:pdf_end_offset])
         tmp.flush()
         subprocess.check_call(["/usr/bin/env", "qpdf", "-qdf", tmp.name, '-'], stdout=FNULL)
     print "PDF appears to be valid."

### 2017-05-10 article.tex
Added a note about it being a polyglot
Added a section on the %%EOF problem

@@
 
+\thindivider
+
+Okay, I have to admit something: There is one more challenge. You see,
+the latest PDF standard, ISO~32000-1:2008, says
+\begin{center}
+\begin{tikzpicture}
+\node at (0,0) (text) {\begin{minipage}{0.8\hsize}\small
+The \emph{trailer} of a PDF file enables a conforming reader to
+quickly find the cross-reference table and certain special
+objects. Conforming readers should read a PDF file from its end. The
+last line of the file shall contain only the end-of-file
+marker, \texttt{\%\%EOF}.
+\end{minipage}};
+\node[anchor=north east] at (text.north west) {\LARGE``};
+\node[anchor=south west] at (text.south east) {\LARGE''};
+\end{tikzpicture}
+\end{center}
+Granted, we are producing a PDF that conforms to version~1.4 of the
+specification, which doesn't appear to have that requirement.
+However, at least as early as version 1.3, the specification did have
+an implementation note that Acrobat requires the \texttt{\%\%EOF} to
+be within the last 1024 bytes of the file.  Either way, that's not
+guaranteed to be the case for us, especially since we are moving the
+PDF to be at the beginning of the packfile.  There are almost
+certainly going to be trailing bytes after the PDF's \texttt{\%\%EOF},
+and if the git repository is large, those are likely to be more than
+1024 bytes.
+
+Fortunately, most common PDF readers don't seem to care how many
+trailing bytes there are, at least when the PDF is version~1.4.
+Unfortunately, some readers such as Adobe's try to be ``helpful,''
+silently ``fixing'' the problem and offering to save the fixed version
+upon exit.  We can at least partially fix the PDF, ensuring that
+the \texttt{\%\%EOF} is exactly 20 bytes of the end of the file, by
+creating a second uncompressed git object as the very end of the
+packfile (right before the final 20~byte SHA-1 checksum).  We could
+then move the trailer from the end of the original PDF at the start of
+the pack to the new git object at the end of the pack.  Finally, we
+could encapsulate the ``middle'' objects of the packfile inside a PDF
+stream object, such that they are ignored by the PDF.  The tricky part
+is that we would have to know how many bytes will be in that
+stream \emph{before} we add the PDF to the git database.  That's
+theoretically possible to do \textit{a priori}, but it'd be very labor
+intensive to pull off.  Furthermore, using this approach will
+completely break the inner PDF that is produced by cloning the
+repository, since its trailer will then be in a separate file.
+Therefore, I chose to live with Adobe's helpfulness and not pursue
+this fix for the PoC.
+
+\thindivider
+
+This PDF is a git bundle containing its \LaTeX\ source, as well as all
+of the code necessary to regenerate this polyglot.
+
 {\fontfamily{jkpvos}\selectfont\begin{center}

### 2017-05-10 article.tex
Off-by-one in the page count

@@
 \begin{minipage}{0.8\hsize}\small
-Thus=---thus=, my fellow-neighbours= and associates= in this= great harvest of our learning, now ripening before our eyes=; thus= it is=, by slow steps= of casual increase, that our knowledge physical, metaphysical, physiological, polemical, nautical, mathematical, \ae{}nigmatical, technical, biographical, romantical, chemical, obstetrical, and polyglottical, with fifty other branches= of it, (most of 'em ending as= these do, in ical) have for these four last centuries= and more, gradually been creeping upwards= towards= that Akme of their perfections=, from which, if we may form a conjecture from the advances= of these last \pgfmathparse{int(\thepage + 1)}\pgfmathresult~pages=, we cannot possibly be far off.
+Thus=---thus=, my fellow-neighbours= and associates= in this= great harvest of our learning, now ripening before our eyes=; thus= it is=, by slow steps= of casual increase, that our knowledge physical, metaphysical, physiological, polemical, nautical, mathematical, \ae{}nigmatical, technical, biographical, romantical, chemical, obstetrical, and polyglottical, with fifty other branches= of it, (most of 'em ending as= these do, in ical) have for these four last centuries= and more, gradually been creeping upwards= towards= that Akme of their perfections=, from which, if we may form a conjecture from the advances= of these last \thepage~pages=, we cannot possibly be far off.
 \end{minipage}

### 2017-05-10 article.tex
Cleaned up the ISO standard code
Forgot to syntax color `return`!

@@
 \verb|        reference += |{\color{gray}\verb|(|}\verb|1 << |{\color{gray}\verb|(|}\verb|7 * |{\color{gray}\verb|(|}\verb|bytes_read - 1|{\color{gray}\verb|)))|}\\
-\verb|    return reference, bytes_read|\\
+{\color{blue}\verb|    return|}\verb| reference, bytes_read|\\
 \end{minipage}
@@
 Okay, I have to admit something: There is one more challenge. You see,
-the latest PDF standard, ISO~32000-1:2008, says
+the PDF standard~(ISO~32000-1) says
 \begin{center}

### 2017-05-10 article.tex
Fixed a typo and cleaned up some wording

@@
 that the PDF occur as early in the packfile as possible. At first, I
-considered applying another patch directly the git source code to make
-the uncompressed object first in the packfile. This approach proved to
-be very involved, in part due to git's UNIX design philosophy and
-architecture of generic code reuse. We're already updating the
-packfile's SHA-1 hash due to changing the DEFLATE headers, so instead
-I decided to simply reorder the objects after-the-fact, subsequent to
-the DEFLATE header fix but before we update the hash. The only
-challenge is that moving objects in the packfile has the potential to
-break offset delta objects, since they refer to their base objects via
-a byte offset within the packfile.  Moving the PDF to the beginning
-will break any offset delta objects that occur after the original
-position of the PDF that refer to base objects that occur before the
-original position of the PDF. That's why I had to dive deeper into the
-rabbit hole of the packfile format to understand the delta object
-headers (as you saw at the end of Section~\ref{sec:GitBundleFormat},
-if you were brave enough to finish it), since I originally attempted
-to rewrite the broken ones. Rewriting the broken offset delta objects
-is the \emph{correct} solution, but, in the end, I discovered a much
+considered applying another patch directly to the git source code to
+make the uncompressed object first in the packfile. This approach
+proved to be very involved, in part due to git's UNIX design
+philosophy and architecture of generic code reuse. We're already
+updating the packfile's SHA-1 hash due to changing the DEFLATE
+headers, so instead I decided to simply reorder the objects
+after-the-fact, subsequent to the DEFLATE header fix but before we
+update the hash. The only challenge is that moving objects in the
+packfile has the potential to break offset delta objects, since they
+refer to their base objects via a byte offset within the packfile.
+Moving the PDF to the beginning will break any offset delta objects
+that occur after the original position of the PDF that refer to base
+objects that occur before the original position of the PDF. I
+originally attempted to rewrite the broken offset delta objects, which
+is why I had to dive deeper into the rabbit hole of the packfile
+format to understand the delta object headers (as you saw at the end
+of Section~\ref{sec:GitBundleFormat}, if you were brave enough to
+finish it). Rewriting the broken offset delta objects is
+the \emph{correct} solution, but, in the end, I discovered a much
 simpler way.

### 2017-05-10 article.tex
prevented an unwanted en-dash

@@
 the command line arguments that is already passed by default
-is \texttt{--delta-base-offset}. Running \texttt{git help
+is \texttt{-{}-delta-base-offset}. Running \texttt{git help
 pack-objects} reveals the following:
@@
 \end{center}
-So all we need to do is \emph{remove} the \texttt{--delta-base-offset}
+So all we need to do is \emph{remove} the \texttt{-{}-delta-base-offset}
 argument and git will not include any offset delta objects in the

### 2017-05-10 article.tex
Clarified some of the text.

@@
 guaranteed to be the case for us, especially since we are moving the
-PDF to be at the beginning of the packfile.  There are almost
-certainly going to be trailing bytes after the PDF's \texttt{\%\%EOF},
-and if the git repository is large, those are likely to be more than
-1024 bytes.
+PDF to be at the beginning of the packfile.  There are always going to
+be at least 20 trailing bytes after the PDF's \texttt{\%\%EOF} (namely
+the packfile's final SHA-1 checksum), and if the git repository is
+large, there are likely to be more than 1024 bytes.
 

### 2017-05-28 article.tex
Added a suggestion at the end
If you're going to "fake" output, at least make sure it's sorted properly!
Fixed a misspelling

@@
 $ ls
-article.tex
-article.pdf
+article.pdf article.tex
 \end{verbatim}
@@
 \node at (0,0) (text) {\begin{minipage}{0.8\hsize}\small
-As a matter of fact, G-d just questioned my judgement. He said,
+As a matter of fact, G-d just questioned my judgment. He said,
 `Terry, are you worthy to be the man who makes The Temple? If you are,
@@
 This PDF is a git bundle containing its \LaTeX\ source, as well as all
-of the code necessary to regenerate this polyglot.
+of the code necessary to regenerate this polyglot. Clone it to take a
+look at the history of this document and its associated code!
 

### 2017-05-28 article.tex
Fixed another typo

@@
 upon exit.  We can at least partially fix the PDF, ensuring that
-the \texttt{\%\%EOF} is exactly 20 bytes of the end of the file, by
+the \texttt{\%\%EOF} is exactly 20~bytes from the end of the file, by
 creating a second uncompressed git object as the very end of the

### 2017-06-05 article.tex
Added a link to the GitHub repo
Better filename for the polyglot PDF

@@
 \footnotesize\begin{verbatim}
-$ git clone article.pdf foo
+$ git clone PDFGitPolyglot.pdf foo
 Cloning into 'foo'...
@@
 $ ls
-article.pdf article.tex
+PDFGitPolyglot.pdf PDFGitPolyglot.tex
 \end{verbatim}
@@
 \begingroup
-\setbox9=\hbox{\footnotesize\verb|$ git bundle create article_bundle.pdf --all|}
+\setbox9=\hbox{\footnotesize\verb|$ git bundle create PDFGitPolyglot.pdf --all|}
 \begin{minipage}{\wd9}
@@
 $ git commit article.pdf -m "added"
-$ git bundle create polyglot.pdf --all
+$ git bundle create PDFGitPolyglot.pdf --all
 \end{verbatim}
@@
 of the code necessary to regenerate this polyglot. Clone it to take a
-look at the history of this document and its associated code!
+look at the history of this document and its associated code! The code
+is also hosted on
+GitHub\footnote{\texttt{https://github.com/ESultanik/PDFGitPolyglot}}.
 

### 2017-06-05 fix_oversize_pdf.py
Fixed a regex error

@@ -16,9 +16,9 @@ def parse_obj(pdf_content, logger = None):
             bytes_skipped += len(line) + 1
             continue
         after_obj = pdf_content[bytes_skipped+len(line)+1:]
-        m = re.match(r"^\s*<<.*?\/Length\s+(\d+)\s*\n.*?>>.*?stream\n", after_obj, re.MULTILINE | re.DOTALL)
+        m = re.match(r"^\s*<<((?!\n\s*>>\s*\n).)*?\/Length\s+(\d+)\s*\n.*?>>.*?stream\n", after_obj, re.MULTILINE | re.DOTALL)
         if m:
-            bytes_up_to_endstream = len(m.group(0)) + int(m.group(1))
+            bytes_up_to_endstream = len(m.group(0)) + int(m.group(2))
             after_stream = after_obj[bytes_up_to_endstream:]
             m2 = re.match(r".*?\n*endstream\s*\nendobj\s*\n", after_stream, re.MULTILINE | re.DOTALL)
             if not m2:
@@ -82,6 +82,7 @@ def fix_pdf(pdf_content, output = None, logger = None):
         start, length = parse_obj(pdf_content[offset:], logger = logger)
         if start is None:
             break
+        logger("PDF object starting at %d and ending at %d\n" % (offset + start, offset + start + length))
         objects.append((offset + start, length))
         if length > 0xFFFF - 5 - 4:
             raise Exception("The object at PDF offset %d is %d bytes, which is more than the maximum of %d! This PDF cannot be fixed." % (offset + start, length, 0xFFFF - 5 - 4))

### 2017-06-05 fix_oversize_pdf.py
Fixed an infinite loop if we decide to put a DEFLATE header before the first object

@@ -43,7 +43,7 @@ def calculate_deflate_locations(objects, last_location = 0):
         else:
             next_location = objects[i+1][0]
         if next_location - last_location + 9 > 0xFFFF:
-            return [i] + map(lambda j : j+i, calculate_deflate_locations(objects[i:], last_location = obj[0]))
+            return [i] + map(lambda j : j+i+1, calculate_deflate_locations(objects[i+1:], last_location = obj[0]))
     return []
 
 PDF_HEADER = r"%PDF-1.\d\s*\n%\xD0\xD4\xC5\xD8\s*\n"

### 2017-06-05 fix_oversize_pdf.py
Fixed some more regex parsing errors

@@ -11,11 +11,14 @@ def parse_obj(pdf_content, logger = None):
         if line.startswith('%'):
             bytes_skipped += len(line) + 1
             continue
-        m = re.match(r"^\s*\d+\s+\d+\s+obj$", line)
+        m = re.match(r"^(\s*\d+\s+\d+\s+obj\s*)(<<|$)", line)
         if not m:
-            bytes_skipped += len(line) + 1
+            bytes_skipped += len(line) + 1            
             continue
-        after_obj = pdf_content[bytes_skipped+len(line)+1:]
+        if m.group(2):
+            after_obj = pdf_content[bytes_skipped + len(m.group(1)) + 1:]
+        else:
+            after_obj = pdf_content[bytes_skipped+len(line)+1:]
         m = re.match(r"^\s*<<((?!\n\s*>>\s*\n).)*?\/Length\s+(\d+)\s*\n.*?>>.*?stream\n", after_obj, re.MULTILINE | re.DOTALL)
         if m:
             bytes_up_to_endstream = len(m.group(0)) + int(m.group(2))

### 2017-06-05 make_polyglot.sh
Make sure we don't stage the PDF polyglot when creating the polyglot

@@ -53,6 +53,8 @@ DIR="$( cd "$( dirname "${BASH_SOURCE[0]}" )" && pwd )"
 PATH=$DIR/git:$PATH git bundle create ${OUTPUT}.bundle --do-not-compress `git hash-object $OUTPUT` --all 1>>${OUTPUT}.log 2>&1
 echo "Switching back to branch ${CURRENT_BRANCH}..."
 git checkout $CURRENT_BRANCH 1>>${OUTPUT}.log 2>&1
+echo "Deleting ${OUTPUT} from being staged..."
+git rm --cached -f ${OUTPUT}
 echo "Deleting temporary branch ${BRANCH_NAME}..."
 git branch -D $BRANCH_NAME 1>>${OUTPUT}.log 2>&1
 mv ${OUTPUT}.bundle $OUTPUT

### 2017-06-06 make_polyglot.sh
Portability improvements

@@ -1,3 +1,5 @@
+#!/usr/bin/env bash
+
 if [ "$#" -ne 2 ]; then
     (>&2 echo "Usage: $0 SOURCE_PDF OUTPUT_PDF\n")
     exit 1
@@ -54,7 +56,8 @@ PATH=$DIR/git:$PATH git bundle create ${OUTPUT}.bundle --do-not-compress `git ha
 echo "Switching back to branch ${CURRENT_BRANCH}..."
 git checkout $CURRENT_BRANCH 1>>${OUTPUT}.log 2>&1
 echo "Deleting ${OUTPUT} from being staged..."
-git rm --cached -f ${OUTPUT}
+echo "It is normal for the following command to produce an error:" >> ${OUTPUT}.log
+git rm --cached -f ${OUTPUT} 1>>${OUTPUT}.log 2>&1
 echo "Deleting temporary branch ${BRANCH_NAME}..."
 git branch -D $BRANCH_NAME 1>>${OUTPUT}.log 2>&1
 mv ${OUTPUT}.bundle $OUTPUT

### 2017-06-06 fix_oversize_pdf.py
Fixed off-by-one error when there is no newline after `obj`

@@ -16,7 +16,7 @@ def parse_obj(pdf_content, logger = None):
             bytes_skipped += len(line) + 1            
             continue
         if m.group(2):
-            after_obj = pdf_content[bytes_skipped + len(m.group(1)) + 1:]
+            after_obj = pdf_content[bytes_skipped+len(m.group(1)):]
         else:
             after_obj = pdf_content[bytes_skipped+len(line)+1:]
         m = re.match(r"^\s*<<((?!\n\s*>>\s*\n).)*?\/Length\s+(\d+)\s*\n.*?>>.*?stream\n", after_obj, re.MULTILINE | re.DOTALL)

### 2017-06-06 make_polyglot.sh
Added some checks to make sure git won't silelently fail

@@ -5,6 +5,17 @@ if [ "$#" -ne 2 ]; then
     exit 1
 fi
 
+# Make sure that the git username and E-mail variables are set,
+# else some of the commands below will fail:
+if [[ -z `git config --list | grep user.email | sed 's/^[^=]*=//'` ]]; then
+    (>&2 echo -e "\nYou do not have the git user.email variable set!\nPlease run \`git config --global user.email \"your@email.com\"\`\n")
+    exit 2
+fi
+if [[ -z `git config --list | grep user.name | sed 's/^[^=]*=//'` ]]; then
+    (>&2 echo -e "\nYou do not have the git user.name variable set!\nPlease run \`git config --global user.name \"Your Name\"\`\n")
+    exit 3
+fi
+
 SOURCE=$1
 OUTPUT=$2
 

### 2017-06-06 make_polyglot.sh
More informative error message, with workaround suggestion

@@ -27,14 +27,14 @@ BRANCH_NAME=PolyglotBranch
 
 git ls-files $OUTPUT --error-unmatch 1>/dev/null 2>&1
 if [ $? -ne 1 ]; then
-    (>&2 echo "Error: The output file $OUTPUT is already tracked by git!\n       This will not work! Please choose another output file.\n")
-    exit 2
+    (>&2 echo "Error: The output file $OUTPUT is already tracked by git!\n       This will not work! Please choose another output file.\n       This might happen if you just cloned this repo from a PDF polyglot.\n       If so, simply run \`git checkout master && git branch -d $BRANCH_NAME\`\n")
+    exit 4
 fi
 
 git rev-parse --verify $BRANCH_NAME 1>/dev/null 2>&1
 if [ $? -eq 0 ]; then
     (>&2 echo "Error: A branch named $BRANCH_NAME already exists!\n       You need to either delete this branch\n       or set a different value for the BRANCH_NAME variable\n       in $0\n")
-    exit 3
+    exit 5
 fi
 
 CURRENT_BRANCH=`git rev-parse --abbrev-ref HEAD`

### 2017-06-06 make_polyglot.sh
Forgot to pass `-e` to `echo`

@@ -1,7 +1,7 @@
 #!/usr/bin/env bash
 
 if [ "$#" -ne 2 ]; then
-    (>&2 echo "Usage: $0 SOURCE_PDF OUTPUT_PDF\n")
+    (>&2 echo -e "Usage: $0 SOURCE_PDF OUTPUT_PDF\n")
     exit 1
 fi
 
@@ -27,13 +27,13 @@ BRANCH_NAME=PolyglotBranch
 
 git ls-files $OUTPUT --error-unmatch 1>/dev/null 2>&1
 if [ $? -ne 1 ]; then
-    (>&2 echo "Error: The output file $OUTPUT is already tracked by git!\n       This will not work! Please choose another output file.\n       This might happen if you just cloned this repo from a PDF polyglot.\n       If so, simply run \`git checkout master && git branch -d $BRANCH_NAME\`\n")
+    (>&2 echo -e "Error: The output file $OUTPUT is already tracked by git!\n       This will not work! Please choose another output file.\n       This might happen if you just cloned this repo from a PDF polyglot.\n       If so, simply run \`git checkout master && git branch -d $BRANCH_NAME\`\n")
     exit 4
 fi
 
 git rev-parse --verify $BRANCH_NAME 1>/dev/null 2>&1
 if [ $? -eq 0 ]; then
-    (>&2 echo "Error: A branch named $BRANCH_NAME already exists!\n       You need to either delete this branch\n       or set a different value for the BRANCH_NAME variable\n       in $0\n")
+    (>&2 echo -e "Error: A branch named $BRANCH_NAME already exists!\n       You need to either delete this branch\n       or set a different value for the BRANCH_NAME variable\n       in $0\n")
     exit 5
 fi
 
